---
title: "Literature Search Table"
author: "Guy Mercer"
date: "23/08/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
lit_search_table <- read.csv(file = "./input/study_pure_or_formulation.csv", header = TRUE)

colnames(lit_search_table) <- c("Insecticide", "Search Term", "Search Type", "Database", "Studies Using Formulation (All)", "Studies Using Formulation (Artificial Exposure)")

database_overview_table <- read.csv(file = "./input/database_overview_table.csv")

colnames(database_overview_table) <- c("Insecticide", "Formulation", "Countries Approved in for Outdoor Application to Bee Attractive Crops (24)", " Countries Without Full Flowering Window Restrictions for All Crops (24)", "Outdoor Applications to Bee Attractive Crops Approved", "Crops Without Full Flowering Window Restrictions")
```

How to use flextable resources:

[Book](https://ardata-fr.github.io/flextable-book/)
[Function Reference](https://davidgohel.github.io/flextable/reference/index.html)

and a reference on colour choices in R:

[R Colours](https://www.datanovia.com/en/blog/awesome-list-of-657-r-color-names/)

Literature Search Table First 

```{r, echo=FALSE}
library(flextable)
library(scales)
library(webshot)

table <- flextable(head(lit_search_table))

table <- add_header_row(table, values = c("Insecticide Literature Searches"), colwidths = c(6))

table <- fontsize(table, i = 1, j = NULL, size = 18, part = "header")

table <- bold(table, i = 1, j = NULL, bold = TRUE, part = "header")

table <- fontsize(table, i = c(1,2), j = 1, size = 12, part = "body")

table <- fontsize(table, i = c(1,2), j = 2, size = 8, part = "body")

table <- bold(table, i = c(1,2), j = c(5,6), bold = TRUE, part = "body")

table <- align(table, i = 1, j = NULL, align = "center", part = "header")

table <- hrule(table, i = 1, rule = "exact", part = "header")

table <- height(table, i = 1, 0.6, part = "header")

table <- width(table, j = 2, 3)

table <- width(table, j = 5, 2)

table <- width(table, j = 6, 2.5)

table <- font(
  table,
  i = NULL,
  j = NULL,
  "calibri",
  part = "all",
)

# background colour

# use scales package to set transparency
colour <- alpha("#999999", alpha = 0.1)

colour_1 <- alpha("#E1BE6A", alpha = 0.4)

colour_2 <- alpha("#40B0A6", alpha = 0.4)


table <- bg(table, i = NULL, j = NULL, bg = colour, part = "body")

table <- bg(table, i = 2, j = NULL, bg = colour, part = "header")

table <- bg(table, i = 1, j = NULL, bg = colour_1, part = "body")

table <- bg(table, i = 2, j = NULL, bg = colour_2, part = "body")

table <- hline(table, i = 1, j = c(2:6), border = NULL, part = "body")

table

# save_as_image(table, path = "./literature_search_table.png")
```

Database Table Second

```{r, echo=FALSE}
table_1 <- flextable(database_overview_table)

table_1 <- add_header_row(table_1, values = c("EU Database Search Overview"), colwidths = c(6))

table_1 <- fontsize(table_1, i = 1, j = NULL, size = 18, part = "header")

table_1 <- fontsize(table_1, i = 2, j = NULL, size = 8, part = "header")

table_1 <- fontsize(table_1, i = 2, j = c(1,2), size = 10, part = "header")

table_1 <- fontsize(table_1, i = NULL, j = 2, size = 10, part = "body")

table_1 <- bold(table_1, i = 1, j = NULL, bold = TRUE, part = "header")

table_1 <- fontsize(table_1, i = c(1:6), j = 1, size = 12, part = "body")

table_1 <- bold(table_1, i = NULL, j = c(4,6), bold = TRUE, part = "body")

table_1 <- align(table_1, i = 1, j = NULL, align = "center", part = "header")

table_1 <- hrule(table_1, i = 1, rule = "exact", part = "header")

table_1 <- height(table_1, i = 1, 0.6, part = "header")

table_1 <- width(table_1, j = c(2:6), 3)

# use scales package to set transparency

colour <- alpha("#999999", alpha = 0.1)

table_1 <- bg(table_1, i = 2, j = NULL, bg = colour, part = "header")

table_1 <- hline(table_1, i = 4, j = c(2:6), border = NULL, part = "body")

table_1 <- hline(table_1, i = 5, j = c(2:6), border = NULL, part = "body")

table_1 <- bold(table_1, i = 2, j = c(4,6), bold = TRUE, part = "header")

table_1 <- align(table_1, i= 2, j = c(2:6), align = "left", part = "header")

table_1 <- align(table_1, i= NULL, j = c(2:6), align = "center", part = "body")

table_1

save_as_image(table_1, path = "./database_overview_table.png", webshot = "webshot2")

```

```{r}
database_overview_table_aggregated <- read.csv("./input/database_overview_table_aggregated.csv")

colnames(database_overview_table_aggregated) <- c("Insecticide", "Countries Without Full Flowering Window Restrictions for All Crops (24)", "Crops Without Full Flowering Window Restrictions")

table_2 <- flextable(database_overview_table_aggregated)

table_2 <- add_header_row(table_2, values = c("EU Database Search Overview"), colwidths = c(3))

table_2 <- fontsize(table_2, i = 1, j = NULL, size = 18, part = "header")

table_2 <- fontsize(table_2, i = 2, j = NULL, size = 8, part = "header")

table_2 <- fontsize(table_2, i = 2, j = 1, size = 10, part = "header")

table_2 <- fontsize(table_2, i = NULL, j = 2, size = 10, part = "body")

table_2 <- bold(table_2, i = 1, j = NULL, bold = TRUE, part = "header")

table_2 <- fontsize(table_2, i = c(1:3), j = 1, size = 12, part = "body")

table_2 <- align(table_2, i = 1, j = NULL, align = "center", part = "header")

table_2 <- hrule(table_2, i = 1, rule = "exact", part = "header")

table_2 <- height(table_2, i = 1, 0.6, part = "header")

table_2 <- width(table_2, j = c(2:3), 3)

# use scales package to set transparency

colour <- alpha("#999999", alpha = 0.1)

table_2 <- bg(table_2, i = 2, j = NULL, bg = colour, part = "header")

table_2 <- hline(table_2, i = 1, j = c(2:3), border = NULL, part = "body")

table_2 <- hline(table_2, i = 2, j = c(2:3), border = NULL, part = "body")

table_2 <- align(table_2, i= 2, j = c(2:3), align = "left", part = "header")

table_2 <- align(table_2, i= NULL, j = c(2:3), align = "center", part = "body")

table_2

save_as_image(table_2, path = "./database_overview_table_aggregated.png", webshot = "webshot2")

```

Now prepare the approved_eu_application_regime table. I am going to import the csv with all the approved EU crop applications minus Bulgaria as there website was down. I'm then going to filter out the crop applications that state no application during flowering window. From the rest I am going to calculate the median dosage regime in g a.s./ha and return that for each country to demonstrate what an average application regime is for each country. 

```{r}
library(tidyverse)

approved_eu_application_regimes <- read.csv("./input/approved_eu_application_regimes.csv")

# remove slovakia, slovenia and apple/pear for italy
approved_eu_application_regimes <- approved_eu_application_regimes [-grep("Slovakia|Slovenia", approved_eu_application_regimes$Country), ]


# can't think of a quick way to search for both Italy and pear/apple in the same function as they are in different columns
# this a poor long winded solution.
italy_row <- grep("Italy", approved_eu_application_regimes$Country)

pear_row <- grep("pear", approved_eu_application_regimes$Crop)

matching_row <- tibble()

for (i in 1:length(italy_row)) {
  
  for (j in 1:length(pear_row)) {
    
   if (italy_row [i] == pear_row [j]) {
     
     matching_row <- rbind(matching_row,italy_row [i])
     
   }
    
  }
  
}

italy_row <- grep("Italy", approved_eu_application_regimes$Country)

apple_row <- grep("apple", approved_eu_application_regimes$Crop)

for (i in 1:length(italy_row)) {
  
  for (j in 1:length(apple_row)) {
    
   if (italy_row [i] == apple_row [j]) {
     
     matching_row <- rbind(matching_row,italy_row [i])
     
   }
    
  }
  
}

approved_eu_application_regimes <- approved_eu_application_regimes [-c(matching_row [1,1], matching_row [2,1]), ]

# remove Spain tobacco as there are ? 
approved_eu_application_regimes <- approved_eu_application_regimes [apply(approved_eu_application_regimes != "?", 1, all), ]

# calculate the median dosage per application for each country
colnames(approved_eu_application_regimes) [colnames(approved_eu_application_regimes) == "Mass.of.Active.Substance.Per.Application..g.a.s..ha."] <- "mass_active_sub"

# sort class
approved_eu_application_regimes$mass_active_sub <- as.numeric(as.character(approved_eu_application_regimes$mass_active_sub))

# remove regimes where there is more than one a year
approved_eu_application_regimes <- approved_eu_application_regimes [approved_eu_application_regimes$Application.Frequency != "4/year",]

median_regimes_df <- tibble()

for (i in 1:length(unique(approved_eu_application_regimes$Country))) {
  
  by_country <- approved_eu_application_regimes [approved_eu_application_regimes$Country == unique(approved_eu_application_regimes$Country) [i], ]
  
  median_mass <- median(by_country$mass_active_sub)
  
  closest_regimes_index <- which(abs(by_country$mass_active_sub - median_mass) == min(abs(by_country$mass_active_sub - median_mass)))
  
  # randomly select one
  closest_regimes_index <- sample(closest_regimes_index, 1)
  
  median_regimes <- by_country [closest_regimes_index, ]
  
  median_regimes_df <- rbind(median_regimes_df, median_regimes)
  
}

colnames(median_regimes_df) <- c("Country", "Delivery Method", "Concentration (g/l)", "Crop", "Application Volume (l/ha)", "Application Frequency", "Mass of Active Substance per Application (g a.s./ha)")
```

Now make this table look nice

```{r}
library(flextable)

table_3 <- flextable(median_regimes_df)

table_3 <- add_header_row(table_3, values = c("Median Application Regime By Country For Sivanto Prime"), colwidths = c(7))

table_3 <- fontsize(table_3, i = 1, j = NULL, size = 18, part = "header")

table_3 <- fontsize(table_3, i = 2, j = NULL, size = 8, part = "header")

table_3 <- bold(table_3, i = 1, j = NULL, bold = TRUE, part = "header")

table_3 <- align(table_3, i = 1, j = NULL, align = "center", part = "header")

table_3 <- hrule(table_3, i = 1, rule = "exact", part = "header")

table_3 <- bold(table_3, i = NULL, j = 7, bold = TRUE, part = "body")

table_3 <- width(table_3, j = 7, 2)

# use scales package to set transparency
colour <- alpha("#999999", alpha = 0.1)

table_3 <- bg(table_3, i = 2, j = NULL, bg = colour, part = "header")

table_3

save_as_image(table_3, path = "./median_application_regimes_by_country.png", webshot = "webshot2")

```

Table for the degradation datasets

```{r}
degradation_datasets <- read.csv("./input/dar_degradation_datasets.csv")

colnames(degradation_datasets) <- c("Study ID", "Crop", "Application Rate", "Measurement Window Start Nectar", "Measurement Window End Nectar", "Timepoints Nectar",
                                    "Measurement Window Start Pollen", "Measurement Window End Pollen", "Timepoints Pollen")

# remove columns 
degradation_datasets <- degradation_datasets [, grep("Study ID|Crop|Application Rate|Timepoints Nectar|Timepoints Pollen", colnames(degradation_datasets))]

library(flextable)

table_4 <- flextable(degradation_datasets)

table_4 <- add_header_row(table_4, values = c("Public Degradation Datasets For Sivanto Prime"), colwidths = c(5))

table_4 <- fontsize(table_4, i = 1, j = NULL, size = 18, part = "header")

table_4 <- fontsize(table_4, i = 2, j = NULL, size = 8, part = "header")

table_4 <- fontsize(table_4, i = NULL, j = NULL, size = 8, part = "body")

table_4 <- fontsize(table_4, i = NULL, j = c(1,3), size = 10, part = "body")

table_4 <- bold(table_4, i = 1, j = NULL, bold = TRUE, part = "header")

table_4 <- align(table_4, i = 1, j = NULL, align = "center", part = "header")

table_4 <- hrule(table_4, i = 1, rule = "exact", part = "header")

table_4 <- bold(table_4, i = NULL, j = 3, bold = TRUE, part = "body")

table_4 <- width(table_4, j = 3, 5)

# use scales package to set transparency
colour <- alpha("#999999", alpha = 0.1)

table_4 <- bg(table_4, i = 2, j = NULL, bg = colour, part = "header")

table_4

save_as_image(table_4, path = "./dar_degradation_datasets.png", webshot = "webshot2")
```

