---
title: "Cake Based Plots"
author: "Guy Mercer"
date: "30/09/2021"
output: html_document
---
Input pollen and nectar data removing points before 0. 

```{r}
nectar_data <- read.csv(file ="input/nectar_residues_ugkg.csv")

pollen_data <- read.csv(file ="input/pollen_residues_ugkg.csv")

nectar_data <- nectar_data [nectar_data$DALA_h > 0, ]

pollen_data <- pollen_data [pollen_data$DALA_h > 0, ]

```

First write the equation and return the line fitted by CAKE using the parameter estimates CAKE outputted. Do this for nectar first. Nectar the best fit was a bi-exponential model, also known as a Double First Order in Parallel model (DFOP). This has 4 parameters - Parent_0, k1_Parent, k2_Parent and g_Parent

```{r}
Parent_0 <- 4160

k1_Parent <- 0.4179

k2_Parent <- 0.02075

g_Parent <- 0.7392

x <- 1

nectar_dfop_line <- function(x) {Parent_0 * (g_Parent * exp(-k1_Parent * x) + (1 - g_Parent) * exp(-k2_Parent * x))}

# halfway between 0-24
nectar_dfop_line(x = 12)
# halfway between 24-48
nectar_dfop_line(x = 36)
# halfway between 48-96
nectar_dfop_line(x = 72)
# halfway between 96-144
nectar_dfop_line(x = 120)
# halfway between 144-192
nectar_dfop_line(x = 168)
# halfway between 192-240
nectar_dfop_line(x = 216)
```

Now do the same for pollen. For pollen the only fit with confident parameters was the SFO..

```{r}
Parent_0 <- 74300
k_Parent <- 0.1225

  
sfo_line <- function(x) {Parent_0 * exp(-k_Parent * x)}

sfo_line(x = 12)
sfo_line(x = 36)
```

pollen fomc

```{r}
Parent_0 <- 80600
alpha <- 1.809
beta <- 9.91

fomc_line <- function(x) {Parent_0 / ((x / beta) + 1) ^ alpha}

fomc_line(x = 12)
fomc_line(x = 36)
mean(pollen_data$residue_measurement_ugkg [pollen_data$DALA_h > 24])
```

Visualise nectar in ggplot.

```{r}
library(tidyverse)

# rename datframe
nectar_total <- nectar_data
colnames(nectar_total) <- c("t", "parent")

# parameter estimates
Parent_0 <- 4160

k1_Parent <- 0.4179

k2_Parent <- 0.02075

g_Parent <- 0.7392

x <- 1

# model function
nectar_dfop_line <- function(x) {Parent_0 * (g_Parent * exp(-k1_Parent * x) + (1 - g_Parent) * exp(-k2_Parent * x))}

# degradation model plot function

# rwrap function allows an interval of x to be defined to draw the model fit line
rwrap=function(f,xmin,xmax){ff=function(x){y=f(x);y[x>xmax]=NA;y[x<xmin]=NA;y}}

# function
degradation_plot_function <- function(dataframe, model_prediction, input_colour, x_min_line, title) {
  
ggplot(data = dataframe, aes(x = t, y = parent)) +
    
xlab("Time (h)") +
    
ylab("Residue Measurements (µg/kg)") +

geom_point(color = input_colour, alpha = 0.7) +

stat_function(fun = rwrap(model_prediction, x_min_line, 240), colour = input_colour, alpha = 0.7) +

theme_bw() +
    
ggtitle(title) +
  
theme(legend.position = "none", axis.title = element_text(size=8, face = "bold"), plot.title = element_text(hjust = 0.5, size = 14, face = "bold")) +
    
scale_x_continuous(limits = c(0, NA), breaks=seq(0,240,24)) # starts axis at 0 and defines ticks

}

# colour
colour <- "#E66100"
colour_lines <- "#5D3A9B"

nectar_total_plot <- degradation_plot_function(dataframe = nectar_total, model_prediction = nectar_dfop_line, input_colour = colour, x_min_line = 0, title = "Nectar Dosage Regime")

# adding the lines. 

# define y values
# halfway between 0-24
t0_24 <- nectar_dfop_line(x = 12)
# halfway between 24-48
t24_48 <- nectar_dfop_line(x = 36)
# halfway between 48-96
t48_96 <- nectar_dfop_line(x = 72)
# halfway between 96-144
t96_144 <- nectar_dfop_line(x = 120)
# halfway between 144-192
t144_192 <- nectar_dfop_line(x = 168)
# halfway between 192-240
t192_240 <- nectar_dfop_line(x = 216)

# add lines
nectar_total_plot <- nectar_total_plot + geom_segment(aes(x=0,xend=24,y=t0_24,yend=t0_24), inherit.aes = FALSE, color = colour_lines, alpha = 0.7, data = data.frame()) # this weird solution is because the segment is drawn nrow(dataframe) times. What is done here is using an empty                                       dataframe to get round this unwanted behaviour link:                                                                                                 (https://stackoverflow.com/questions/48397603/alpha-in-geom-segment-not-working). 

nectar_total_plot <- nectar_total_plot + geom_segment(aes(x=24,xend=48,y=t24_48,yend=t24_48), inherit.aes = FALSE, color = colour_lines, alpha = 0.7, data = data.frame())

nectar_total_plot <- nectar_total_plot + geom_segment(aes(x=48,xend=96,y=t48_96,yend=t48_96), inherit.aes = FALSE, color = colour_lines, alpha = 0.7, data = data.frame())

nectar_total_plot <- nectar_total_plot + geom_segment(aes(x=96,xend=144,y=t96_144,yend=t96_144), inherit.aes = FALSE, color = colour_lines, alpha = 0.7, data = data.frame())

nectar_total_plot <- nectar_total_plot + geom_segment(aes(x=144,xend=192,y=t144_192,yend=t144_192), inherit.aes = FALSE, color = colour_lines, alpha = 0.7, data = data.frame())

nectar_total_plot <- nectar_total_plot + geom_segment(aes(x=192,xend=240,y=t192_240,yend=t192_240), inherit.aes = FALSE, color = colour_lines, alpha = 0.7, data = data.frame())


nectar_total_plot <- nectar_total_plot + geom_segment(aes(x=24,xend=24,y=t0_24,yend=t24_48), inherit.aes = FALSE, color = colour_lines, alpha = 0.7, data = data.frame())

nectar_total_plot <- nectar_total_plot + geom_segment(aes(x=48,xend=48,y=t24_48,yend=t48_96), inherit.aes = FALSE, color = colour_lines, alpha = 0.7, data = data.frame())

nectar_total_plot <- nectar_total_plot + geom_segment(aes(x=96,xend=96,y=t48_96,yend=t96_144), inherit.aes = FALSE, color = colour_lines, alpha = 0.7, data = data.frame())

nectar_total_plot <- nectar_total_plot + geom_segment(aes(x=144,xend=144,y=t96_144,yend=t144_192), inherit.aes = FALSE, color = colour_lines, alpha = 0.7, data = data.frame())

nectar_total_plot <- nectar_total_plot + geom_segment(aes(x=192,xend=192,y=t144_192,yend=t192_240), inherit.aes = FALSE, color = colour_lines, alpha = 0.7, data = data.frame())

nectar_total_plot

# save
 cairo_pdf(file = "./output/nectar_degradation.pdf",   # The directory you want to save the file in
     width = 8, # The width of the plot in inches
     height = 6) # The height of the plot in inches

nectar_total_plot

 dev.off()
```

Visualise pollen in ggplot

```{r}
# rename dataframe
pollen_total <- pollen_data
colnames(pollen_total) <- c("t", "parent")

# parameter estimates
Parent_0 <- 80600

alpha <- 1.809

beta <- 9.91

fomc_line <- function(x) {Parent_0 / ((x / beta) + 1) ^ alpha}

# model function
pollen_fomc_line <- function(x) {Parent_0 / ((x / beta) + 1) ^ alpha}

# colour
colour <- "#1A85FF"
colour_lines <- "#D41159"

pollen_total_plot <- degradation_plot_function(dataframe = pollen_total, model_prediction = pollen_fomc_line, input_colour = colour, x_min_line = 0, title = "Pollen Dosage Regime")

# adding the lines. 

# define y values

# halfway between 0-24
t0_24 <- fomc_line(x = 12)
# halfway between 24-48
t24_48 <- fomc_line(x = 36)
# from 48-240
t48_240 <- mean(pollen_data$residue_measurement_ugkg [pollen_data$DALA_h > 24])

# add lines
pollen_total_plot <- pollen_total_plot + geom_segment(aes(x=0,xend=24,y=t0_24,yend=t0_24), inherit.aes = FALSE, color = colour_lines, alpha = 0.7, data = data.frame()) # this weird solution is because the segment is drawn nrow(dataframe) times. What is done here is using an empty                                       dataframe to get round this unwanted behaviour link:                                                                                                 (https://stackoverflow.com/questions/48397603/alpha-in-geom-segment-not-working). 

pollen_total_plot <- pollen_total_plot + geom_segment(aes(x=24,xend=48,y=t24_48,yend=t24_48), inherit.aes = FALSE, color = colour_lines, alpha = 0.7, data = data.frame())

pollen_total_plot <- pollen_total_plot + geom_segment(aes(x=48,xend=240,y=t48_240,yend=t48_240), inherit.aes = FALSE, color = colour_lines, alpha = 0.7, data = data.frame())


pollen_total_plot <- pollen_total_plot + geom_segment(aes(x=24,xend=24,y=t0_24,yend=t24_48), inherit.aes = FALSE, color = colour_lines, alpha = 0.7, data = data.frame())

pollen_total_plot <- pollen_total_plot + geom_segment(aes(x=48,xend=48,y=t24_48,yend=t48_240), inherit.aes = FALSE, color = colour_lines, alpha = 0.7, data = data.frame())

pollen_total_plot

cairo_pdf(file = "./output/pollen_degradation.pdf",   # The directory you want to save the file in
  width = 8, # The width of the plot in inches
  height = 6) # The height of the plot in inches

pollen_total_plot

 dev.off()
```
