---
title: "Cake Based Plots"
author: "Guy Mercer"
date: "30/09/2021"
output: html_document
---
Input pollen and nectar data removing points before 0. 

```{r}
nectar_data <- read.csv(file ="input/nectar_residues_ugkg.csv")

pollen_data <- read.csv(file ="input/pollen_residues_ugkg.csv")

nectar_data <- nectar_data [nectar_data$DALA_h > 0, ]

pollen_data <- pollen_data [pollen_data$DALA_h > 0, ]

```

First write the equation and return the line fitted by CAKE using the parameter estimates CAKE outputted. Do this for nectar first. Nectar the best fit was a bi-exponential model, also known as a Double First Order in Parallel model (DFOP). This has 4 parameters - Parent_0, k1_Parent, k2_Parent and g_Parent

```{r}
Parent_0 <- 4160

k1_Parent <- 0.4179

k2_Parent <- 0.02075

g_Parent <- 0.7392

x <- 1

nectar_dfop_line <- function(x) {Parent_0 * (g_Parent * exp(-k1_Parent * x) + (1 - g_Parent) * exp(-k2_Parent * x))}

nectar_dfop_line(x = 12)
nectar_dfop_line(x = 36)
nectar_dfop_line(x = 60)
nectar_dfop_line(x = 108)
nectar_dfop_line(x = 156)
nectar_dfop_line(x = 204)




```

Now do the same for pollen. For pollen the only fit with confident parameters was the SFO..

```{r}
Parent_0 <- 74300
k_Parent <- 0.1225

  
sfo_line <- function(x) {Parent_0 * exp(-k_Parent * x)}

sfo_line(x = 12)
sfo_line(x = 36)
```

pollen fomc

```{r}
Parent_0 <- 80600
alpha <- 1.809
beta <- 9.91

fomc_line <- function(x) {Parent_0 / ((x / beta) + 1) ^ alpha}

fomc_line(x = 12)
fomc_line(x = 36)
fomc_line(x = 60)

```

