---
title: "time_to_egg_laying"
author: "Guy Mercer"
date: "04/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
# import csv
egg_laying_df <- read.csv("input/egg-laying-data.csv")

colnames(egg_laying_df) <- c("microcolony", "batch","colony", "treatment", "egg_laying_day", 
                             "day")

# set variables as factors with appropriate reference levels
egg_laying_df$treatment <- as.factor(egg_laying_df$treatment)
egg_laying_df$treatment <- relevel(egg_laying_df$treatment, "control")

egg_laying_df$batch <- as.factor(egg_laying_df$batch)
egg_laying_df$batch <- relevel(egg_laying_df$batch, "1")

egg_laying_df$colony <- as.factor(egg_laying_df$colony)
egg_laying_df$colony <- relevel(egg_laying_df$colony, "1")

egg_laying_df$day <- as.numeric(egg_laying_df$day)

# remove column
egg_laying_df <- egg_laying_df [, -5]

# remove microcolonies that had already laid eggs before exposure began
egg_laying_df <- egg_laying_df [egg_laying_df$day != 0,]

# remove sivanto rows
egg_laying_df <- egg_laying_df [egg_laying_df$treatment != "sivanto",]

egg_laying_df$treatment <- factor(egg_laying_df$treatment )

# add event Y/N column
egg_laying_df$laying_event <- 1
```

```{r}
library(survival)
library(coxme)

fit1 <- survfit(Surv(day, laying_event) ~ batch + treatment, data=egg_laying_df)

print(fit1)

summary(fit1, times = (0:14))

plot(fit1, col=1:6, lwd=2, mark.time=TRUE, xlab="Days Since Exposure", ylab="Egg Laying", #conf.int = TRUE
    fun='event')

legend(7, .5, c("batch=1, treatment=control", "batch=1, treatment=fpf-pure", "batch=2, treatment=control",
                  "batch=2, treatment=fpf-pure", "batch=3, treatment=control", "batch=3, treatment=fpf-pure"),
       col=1:6, lwd=2, bty='n')

# for the sake of visualisation only include treatment
fit2 <- survfit(Surv(day, laying_event) ~ treatment, data=egg_laying_df)

print(fit2)

summary(fit2, times = (0:14))

plot(fit2, col=1:2, lwd=2, mark.time=TRUE, xlab="Days Since Exposure", ylab="Egg Laying", #conf.int = TRUE
    fun='event')

legend(7, .5, c("treatment=control", "treatment=fpf-pure"),
       col=1:2, lwd=2, bty='n')
```

```{r}
cfit1 <- coxme(Surv(day, laying_event) ~ batch * treatment + (1 | colony), data=egg_laying_df)

cfit2 <- update(cfit1, .~. - batch:treatment)
anova(cfit1, cfit2)

cfit3 <- update(cfit2, .~. - batch)
anova(cfit2, cfit3)

cfit4 <- update(cfit2, .~. - treatment)
anova(cfit2, cfit4)

cfit5 <- update(cfit4, .~. - batch)
anova(cfit4, cfit5)

# just treatment
summary(cfit3)
```

```{r}
# proportional hazards
zp1 <- cox.zph(cfit3)
zp1

# plot(zp1[2])
plot(zp1[1])
```

Passes proportionality test. 

Neither batch or treatment had an effect on the hazards ratio for egg laying.  

```{r}
# cfit3 <- coxph(Surv(day, laying_event) ~ treatment, data=egg_laying_df)

# plot(survfit(cfit3))

# newdat <- with(egg_laying_df, 
#                data.frame(
#                  treatment = c("control", "fpf-pure")
#                )
# )

# plot(survfit(cfit3, newdata = newdat), conf.int = TRUE, col = c("red", "blue"), fun='event')
# legend(11, 0.2, 
#        legend=c("control", "treatment"),
#        lty = 1, 
#        col = c("red", "blue"),
#        text.col = c("red", "blue"))
```

The survival function S(t) is the probability of surviving until at least time t. It is essentially estimated by the number of patients alive at time t divided by the total number of study subjects remaining at that time.

The survival function at the second death time, t(2) is equal to the estimated probability of not dying at time t(2) conditional on the individual being still at risk at time t(2). 

```{r}
library(survminer)

# log rank test between KM survival curves.
surv <- Surv(time = egg_laying_df$day, event = egg_laying_df$laying_event)
survdiff(surv ~ egg_laying_df$treatment)

# risk.table = TRUE,  risk.table.height = 0.25

# for prettier graphs
p <- ggsurvplot(fit2,
                data = egg_laying_df,
                fun='event',
                conf.int = TRUE,
                palette = c("#117733", "#332288"),
                legend.labs = c("control", "pesticide"),
                xlab = "Time (days)",
                ylab = "Probability of Egg Laying",
                legend.title = "",
                title = "Time To Egg Laying")


p <- p + theme_survminer(font.main = c(24, "bold", "black"),
                         font.x = c(18, "bold", "black"),
                         font.y = c(18, "bold", "black"),
                         font.tickslab = c(16, "plain", "black"),
                         font.legend = c(16, "plain", "black"))

p <- p$plot + theme(plot.title = element_text(hjust = 0.5, face = "bold")) + annotate(geom = "text", x = 14, y = 0.02, label = "bolditalic('p')~bold('= 0.841')", fontface = 2, size = 6, parse = TRUE)

p
```

```{r}
cairo_pdf(file = "./egg_laying_plot.pdf",   # The directory you want to save the file in
  width = 8, # The width of the plot in inches
  height = 6) # The height of the plot in inches

p

 dev.off()
```

Two sources I used for below info 

[link 1](https://rpubs.com/daspringate/survival)
[link 2](https://thomaselove.github.io/432-notes/cox-regression-models-for-survival-data-example-1.html)

Hazard Function
Survival models can be viewed as consisting of two parts: the underlying hazard function, describing how the risk of event per time unit changes over time at baseline levels of covariates; and the effect parameters, describing how the hazard varies in response to explanatory covariates.

h(t|x) = h0(t) exp[B1.x]

h0(t) = is the baseline hazard, which is a non-parametric and unspecified value which depends on t but not on x.

exp[B1.x] = how the baseline hazard is modified as x changes. 

It is a semi-parametric model:

The baseline hazard function is unspecified
The effects of the covariates are multiplicative (proportional assumption)
Doesn't make arbitrary assumptions about the shape/form of the baseline hazard function

Proportionality Assumption
In a proportional hazards model, the unique effect of a unit increase in a covariate is multiplicative with respect to the hazard rate. For example, taking a drug may halve one’s hazard rate for a stroke occurring, or, changing the material from which a manufactured component is constructed may double its hazard rate for failure.

Covariates multiply the hazard by some constant
e.g. a drug may halve a subjects risk of death at any time
*The effect is the same at any time point*

Accounting for non-proportional hazards
Include an interaction with time for the variables:
This factors time out of the main effect
Only use if it makes sense to have a linear interaction between the covariate and time (look at the graphs!)
