---
title: "pollen_consumption aggregated"
author: "Guy Mercer"
date: "01/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
pollen <- read.csv("output/pollen_consumption_corrected_long.csv")
```

```{r}
library(tidyverse)

# x1000 for mg
pollen$corrected_consumption <- pollen$corrected_consumption * 1000
```

First correct day consumption values for worker number. 

```{r}
worker_number <- read.csv("input/number_of_workers_each_day.csv")

# convert into long format
day <- c("1", "2", "4", "6", "8", "10")

worker_long <- tibble()

for (i in 1:length(day)) {
  
  worker_temp <- worker_number [, c(1:5, 5+i)]
  
  worker_temp$day <- day [i]
  
  colnames(worker_temp) [6] <- "worker_number"
  
  worker_long <- rbind(worker_long, worker_temp)
  
}

# create microcolony_day variable for worker_long and pollen
worker_long$mc_day <- paste0(worker_long$microcolony.number, "_", worker_long$day)

pollen$mc_day <- paste0(pollen$microcolony_number, "_", pollen$day)

# associate the worker number with the pollen dataframe. below says for each row of pollen df
# scan each row of the worker_long df. if there is a match between the mc_day unique identifier
# then put the worker number value for that row of worker_long into the matched row of pollen df
pollen$worker_number <- 0

for (i in 1:nrow(pollen)) {
  
  for (k in 1:nrow(worker_long)) {
    
      if (pollen$mc_day [i] == worker_long$mc_day [k]) {pollen$worker_number [i] <- worker_long$worker_number [k]}
    
  }
}

# remove mc_day from pollen
pollen <- pollen [,-7]

# divide corrected consumption values by number of bees present
pollen$corrected_consumption <- pollen$corrected_consumption / pollen$worker_number
```

Calculate average total consumption per bee per microcolony. May be more accurate to describe this as total pollen processed per bee per microcolony.

```{r}
# sum consumption by MC
tot_con <- tibble()

microcolony_numbers <- unique(pollen$microcolony_number)

for (i in 1:length(microcolony_numbers)) {
  
  if (length(pollen$corrected_consumption [pollen$microcolony_number == microcolony_numbers [i]]) == 6) {

  filtered_con <- cbind(sum(pollen$corrected_consumption [pollen$microcolony_number == microcolony_numbers [i]]), microcolony_numbers [i])
  
  tot_con <- rbind(tot_con, filtered_con)
  
  }

}

# re-associate these MC values with colony, batch etc information
batch_col_info <- unique(pollen [, c(1,2,3,4)])

batch_col_info$corrected_consumption <- 0 

for (i in 1:nrow(tot_con)) {
  
  for (k in 1:nrow(batch_col_info)) {
    
    if (tot_con [i, 2] == batch_col_info [k, 1]) {
      
      batch_col_info$corrected_consumption [k] <- tot_con [i, 1] 
      
    }
    
  }
  
}

# remove 0s (these had spills) and rename data
pollen <- batch_col_info

pollen <- pollen [pollen$corrected_consumption > 0, ]

```

```{r}
# sort out the variable classes
pollen$microcolony_number <- as.factor(pollen$microcolony_number)
pollen$batch <- as.factor(pollen$batch)
pollen$colony_derived_from <- as.factor(pollen$colony_derived_from)
pollen$treatment <- as.factor(pollen$treatment)
```

Now begin analysis. At this point I have corrected the consumption values to account for number of bees present on particular day, summed those values and removed any MCs that didn't have full consumption dataset. 

Bit of exploratory analysis first

```{r}
# Cleveland Dot Plot. First make column with 1 and 2 that correspond to Treatment levels

pollen$clevelandcode <- factor(
ifelse(pollen$treatment == "control",
1,
2))

pollen$clevelandcode <- as.numeric(pollen$clevelandcode)

dotchart(pollen$corrected_consumption,
         groups = factor(pollen$clevelandcode),
         ylab = "Order of observations",
         xlab = "Consumption (mg)", main = "Cleveland dotplot", pch = pollen$clevelandcode)

boxplot(corrected_consumption ~ treatment * batch,
        varwidth = TRUE, xlab = "TreatmentxBatch",
        main = "Boxplot of Consumption Conditional on Treatment and Batch", 
        ylab = "Consumption", data = pollen)

boxplot(corrected_consumption ~ batch,
        varwidth = TRUE, xlab = "Batch",
        main = "Boxplot of Consumption Conditional on Batch", 
        ylab = "Consumption", data = pollen)

boxplot(corrected_consumption ~ treatment,
        varwidth = TRUE, xlab = "Treatment",
        main = "Boxplot of Consumption Conditional on Treatment", 
        ylab = "Consumption", data = pollen)
```

Batch 2 appears to have smaller spread, which is mainly driven by fpf-batch2. 

Let's fit beyond optimal model. 

```{r}
# beyond optimal model.

# biologically this is my beyond optimal model.

M.lm <- lm(corrected_consumption ~ treatment + batch + treatment:batch, data = pollen)

summary(M.lm)

# use the residuals from this beyond optimal model
plot(M.lm, which = c (1))
```

```{r}
library(nlme)

M1.gls <- gls(corrected_consumption ~ treatment + batch + treatment:batch, data = pollen, method = "REML")

E <- resid(M1.gls)

op <- par(mfrow = c(2, 2))

boxplot(E ~ pollen$treatment, main = "Treatment")
abline(0, 0)

boxplot(E ~ pollen$batch, main = "Batch")
abline(0, 0)

boxplot(E ~ pollen$treatment * pollen$batch, main = "Treatment & Batch")
abline(0, 0)

boxplot(E ~ pollen$colony_derived_from, main = "Colony")
abline(0, 0)

par(op)

plot(M1.gls, which = c (1))
```

Colony residual plot suggests colony can be included as a random intercept. Maybe batch in a varIdent structure but doubt it. 

```{r}
# random intercept?
M1.lme <- lme(corrected_consumption ~ treatment + batch + treatment:batch, data = pollen, 
              random = ~1 | colony_derived_from, method = "REML")

AIC(M1.gls, M1.lme)

anova(M1.gls, M1.lme)

library(lmtest)
lrtest(M1.gls, M1.lme)

# What about the variance structure?
M1a.gls <- gls(corrected_consumption ~ treatment + batch + treatment:batch,
               data = pollen,
               method = "REML",
               weights = varIdent(form =~ 1 | treatment*batch))

M1b.gls <- gls(corrected_consumption ~ treatment + batch + treatment:batch,
               data = pollen,
               method = "REML",
               weights = varIdent(form =~ 1 | batch))

AIC(M1.gls, M1a.gls, M1b.gls)

anova(M1.gls, M1b.gls)

anova(M1.gls, M1a.gls)

# no varIdent structure is a significant improvement
```

The above doesn't support including colony as a random effect or batch into the variance structure. Maybe revisit this again during model validation. 

My starting model for model selection is M1.gls

```{r}
M1.gls <- gls(corrected_consumption ~ treatment + batch + treatment:batch, data = pollen, method = "ML")

# what does AIC say?
library(MASS)
stepAIC(M1.gls)
```

Indicates batch is significant. Now manually

```{r}
# manual 

# round 1
M1.gls.a <- update(M1.gls, .~. - treatment:batch)

anova(M1.gls, M1.gls.a)

# round 2 
M1.gls.b <- update(M1.gls.a, .~. - treatment)
M1.gls.c <- update(M1.gls.a, .~. - batch)

anova(M1.gls.a, M1.gls.b)
anova(M1.gls.a, M1.gls.c)

```

Nothing is significant. Leave in treatment and batch as part of expt design. Graphical Model Validation

```{r}
M1.gls.a <- update(M1.gls.a, method = "REML")

# extract residuals
E <- resid(M1.gls.a, type = "normalized")

# residuals vs fitted
plot(M1.gls.a, which = c (1))

op <- par(mfrow = c(2, 2))

boxplot(E ~ pollen$treatment, main = "Treatment")
abline(0, 0)

boxplot(E ~ pollen$batch, main = "Batch")
abline(0, 0)

boxplot(E ~ pollen$treatment * pollen$batch, main = "Treatment & Batch")
abline(0, 0)

boxplot(E ~ pollen$colony_derived_from, main = "Colony")
abline(0, 0)

par(op)
```

The colony residual plot suggests that it should be a random effect. Add and see if it improves the plot.

```{r}
M1.lme.a <- lme(corrected_consumption ~ treatment + batch,
               data = pollen,
               random =~1 | colony_derived_from, method = "REML")

# summary(M1.lme.a)

# extract residuals
E <- resid(M1.lme.a , type = "response")

# residuals vs fitted
plot(M1.lme.a , which = c (1))

op <- par(mfrow = c(2, 2))

boxplot(E ~ pollen$treatment, main = "Treatment")
abline(0, 0)

boxplot(E ~ pollen$batch, main = "Batch")
abline(0, 0)

boxplot(E ~ pollen$treatment * pollen$batch, main = "Treatment & Batch")
abline(0, 0)

boxplot(E ~ pollen$colony_derived_from, main = "Colony")
abline(0, 0)

par(op)

```

residuals by colony are much better so include. Treatment * Batch plot is now not great. Add VarIdent structure.

```{r}
M1.lme.a <- lme(corrected_consumption ~ treatment + batch,
               data = pollen,
               random =~1 | colony_derived_from, method = "REML",
               weights = varIdent(form =~ 1 | batch*treatment))

# summary(M1.lme.a)

# extract residuals
E <- resid(M1.lme.a , type = "response")

# residuals vs fitted
plot(M1.lme.a , which = c (1))

op <- par(mfrow = c(2, 2))

boxplot(E ~ pollen$treatment, main = "Treatment")
abline(0, 0)

boxplot(E ~ pollen$batch, main = "Batch")
abline(0, 0)

boxplot(E ~ pollen$treatment * pollen$batch, main = "Treatment & Batch")
abline(0, 0)

boxplot(E ~ pollen$colony_derived_from, main = "Colony")
abline(0, 0)

par(op)

```

The VarIdent structure didn't help really. Go forward with no VarIdent.

```{r}
M1.lme.a <- lme(corrected_consumption ~ treatment + batch,
               data = pollen,
               random =~1 | colony_derived_from, method = "REML")
```

What are the p-values of treatment and batch

```{r}
M1.lme.int <- lme(corrected_consumption ~ treatment * batch,
                  data = pollen,
                  random =~1 | colony_derived_from, method = "ML")

M1.lme.a <- lme(corrected_consumption ~ treatment + batch,
               data = pollen,
               random =~1 | colony_derived_from, method = "ML")

anova(M1.lme.int, M1.lme.a)

# round 2 
M1.lme.b <- update(M1.lme.a, .~. - treatment)
M1.lme.c <- update(M1.lme.a, .~. - batch)

anova(M1.lme.a, M1.lme.b)
anova(M1.lme.a, M1.lme.c)
```

Extract Predictions and SE for fixed effect component of model. 

```{r}
M1.lme.a <- lme(corrected_consumption ~ treatment + batch,
               data = pollen,
               random =~1 | colony_derived_from, method = "REML")

# summary(M1.lme.a)

newdat <- expand.grid(treatment=levels(pollen$treatment), batch=levels(pollen$batch))

newdat$pred <- predict(M1.lme.a, newdat, level = 0)

Designmat <- model.matrix(eval(eval(M1.lme.a$call$fixed)[-2]), newdat[-3])

predvar <- diag(Designmat %*% M1.lme.a$varFix %*% t(Designmat))

newdat$SE <- sqrt(predvar)

newdat$ci_lower <- newdat$pred - (1.96 * newdat$SE) # 1.96 or qt(1 - (0.05 / 2), n - 1)?

newdat$ci_higher <- newdat$pred + (1.96 * newdat$SE) # 1.96 or qt(1 - (0.05 / 2), n - 1)?
```

```{r}
batch_one_only <- newdat [newdat$batch == "1", ]

# ggplot
p1 <- ggplot(data = batch_one_only, aes(x = treatment, y = pred, colour = as.factor(treatment))) + 
  
  geom_point(position = position_dodge(width = 0.75), size = 3) +
  
  geom_errorbar(aes(x = treatment,
                    ymin = ci_lower,
                    ymax = ci_higher,
                    colour = as.factor(treatment)),
                position = position_dodge(width = 0.75),
                width = 0.05, 
                size = 0.8) +
  
  geom_point(data = pollen,
             aes(x = treatment,
                 y = corrected_consumption,
                 colour = as.factor(treatment)),
             alpha = 0.4,
             position  = position_jitterdodge(),
             show.legend = FALSE) +
  
  theme_bw() +

  labs(y = "Average Pollen Consumption/Bee/Microcolony", colour = "Treatment") +
    
  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    
  ggtitle("Average Pollen Consumption By Treatment") +
    
  theme(plot.title = element_text(hjust = 0.5), axis.title.x = element_blank()) +
  
  scale_color_manual(values = c("#117733", "#332288"), labels = c("Control", "Flupyradifurone")) +
  
  scale_x_discrete(labels=c("Control", "Flupyradifurone")) +
  
  annotate(geom = "text", x = 2.32, y = 145, label = "bolditalic('p')~bold('= 0.344')", fontface = 2, size = 6, parse = TRUE)


p1

```

Now a batch plot for treatment=control

```{r}
control_only <- newdat [newdat$treatment == "control", ]

p <- ggplot(data = control_only, aes(x = batch, y = pred, colour = as.factor(batch))) + 
  
  geom_point(position = position_dodge(width = 0.75), size = 3) +
  
  geom_errorbar(aes(x = batch,
                    ymin = ci_lower,
                    ymax = ci_higher,
                    colour = as.factor(batch)),
                position = position_dodge(width = 0.75),
                width = 0.05, 
                size = 0.8) +
  
  geom_point(data = pollen,
             aes(x = batch,
                 y = corrected_consumption,
                 colour = as.factor(batch)),
             alpha = 0.4,
             position  = position_jitterdodge(),
             show.legend = FALSE) +
  
  theme_bw() +

  labs(y = "Average Pollen Consumption/Bee/Microcolony", colour = "Batch") +
    
  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    
  ggtitle("Average Pollen Consumption By Batch") +
    
  theme(plot.title = element_text(hjust = 0.5), axis.title.x = element_blank()) +
  
  scale_color_manual(values = c("#88CCEE", "#CC6677", "#DDCC77")) +

  annotate(geom = "text", x = 3.25, y = 115, label = "bolditalic('p')~bold('= 0.072')", fontface = 2, size = 6, parse = TRUE)

p
```


```{r}
pdf(file = "./output/pollen_consumption_batch.pdf",   # The directory you want to save the file in
  width = 8, # The width of the plot in inches
  height = 6) # The height of the plot in inches

p

 dev.off()
 
pdf(file = "./output/pollen_consumption_treatment.pdf",   # The directory you want to save the file in
  width = 8, # The width of the plot in inches
  height = 6) # The height of the plot in inches

p1

 dev.off()
```

```{r}

```

