---
title: "sucrose consumption analysis"
author: "Guy Mercer"
date: "15/02/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
# import curated data
sucrose <- read.csv(file = "output/sucrose_consumption_corrected_long.csv", )

pollen <- read.csv(file = "output/pollen_consumption_corrected_long.csv")
```

Start with sucrose. Visualisations.

```{r}
library(lattice)

# sort out the variable classes
sucrose$microcolony_number <- as.factor(sucrose$microcolony_number)
sucrose$batch <- as.factor(sucrose$batch)
sucrose$colony_derived_from <- as.factor(sucrose$colony_derived_from)
sucrose$day <- as.numeric(sucrose$day)

# divide days 4,6,8 and 10 by 2 as they represent consumption over two days

for (i in 1:nrow(sucrose)) {
  
  if (sucrose$day [i] == 4) {sucrose$corrected_consumption [i] <- sucrose$corrected_consumption [i] / 2}
  
  if (sucrose$day [i] == 6) {sucrose$corrected_consumption [i] <- sucrose$corrected_consumption [i] / 2}
  
  if (sucrose$day [i] == 8) {sucrose$corrected_consumption [i] <- sucrose$corrected_consumption [i] / 2}

  if (sucrose$day [i] == 10) {sucrose$corrected_consumption [i] <- sucrose$corrected_consumption [i] / 2}


}

# x1000 for mg
sucrose$corrected_consumption <- sucrose$corrected_consumption * 1000
```

Express in consumption per bee. Shouldn't make much difference as only small number of deaths at the beginning of the experiment.

```{r}
worker_number <- read.csv("input/number_of_workers_each_day.csv")

# convert into long format
day <- c("1", "2", "4", "6", "8", "10")

worker_long <- tibble()

for (i in 1:length(day)) {
  
  worker_temp <- worker_number [, c(1:5, 5+i)]
  
  worker_temp$day <- day [i]
  
  colnames(worker_temp) [6] <- "worker_number"
  
  worker_long <- rbind(worker_long, worker_temp)
  
}

# create microcolony_day variable for worker_long and sucrose
worker_long$mc_day <- paste0(worker_long$microcolony.number, "_", worker_long$day)

sucrose$mc_day <- paste0(sucrose$microcolony_number, "_", sucrose$day)


# associate the worker number with the sucrose dataframe. below says for each row of sucrose df
# scan each row of the worker_long df. if there is a match between the mc_day unique identifier
# then put the worker number value for that row of worker_long into the matched row of sucrose df
sucrose$worker_number <- 0

for (i in 1:nrow(sucrose)) {
  
  for (k in 1:nrow(worker_long)) {
    
      if (sucrose$mc_day [i] == worker_long$mc_day [k]) {sucrose$worker_number [i] <- worker_long$worker_number [k]}
    
  }
}

# remove mc_day from sucrose
sucrose <- sucrose [,-7]

# divide corrected consumption values by number of bees present
sucrose$corrected_consumption <- sucrose$corrected_consumption / sucrose$worker_number
```

Bit of exploratory analysis first

```{r}
# Cleveland Dot Plot. First make column with 1 and 2 that correspond to Treatment levels

sucrose$clevelandcode <- factor(
ifelse(sucrose$treatment == "control",
1,
2))

sucrose$clevelandcode <- as.numeric(sucrose$clevelandcode)

dotchart(sucrose$corrected_consumption,
         groups = factor(sucrose$clevelandcode),
         ylab = "Order of observations",
         xlab = "Consumption (mg)", main = "Cleveland dotplot", pch = sucrose$clevelandcode)

boxplot(corrected_consumption ~ treatment * batch,
        varwidth = TRUE, xlab = "Treatment",
        main = "Boxplot of Consumption Conditional on Treatment and Batch", 
        ylab = "Consumption", data = sucrose)

boxplot(corrected_consumption ~ batch,
        varwidth = TRUE, xlab = "Treatment",
        main = "Boxplot of Consumption Conditional on Batch", 
        ylab = "Consumption", data = sucrose)

boxplot(corrected_consumption ~ treatment,
        varwidth = TRUE, xlab = "Treatment",
        main = "Boxplot of Consumption Conditional on Treatment", 
        ylab = "Consumption", data = sucrose)
```

Four control values that are substantially higher than the rest. 

```{r}
# have a look at the consumption
xyplot(corrected_consumption ~ day, data = sucrose,
       xlab = "Day", ylab = "Sucrose Consumption",
       panel=function(x, y){
         panel.xyplot(x, y, col = 1, cex = 0.5, pch = 1)
         panel.grid(h = -1, v = 2)
         panel.abline(v = 0, lty = 2)
         if (length(x) > 5) panel.loess(x, y, span = 0.6,
                                        col = 1, lwd = 2)
         })

# have a look at the consumption for every microcolony. 
xyplot(corrected_consumption ~ day | microcolony_number, data = sucrose,
       xlab = "Day", ylab = "Sucrose Consumption",
       panel=function(x, y){
         panel.xyplot(x, y, col = 1, cex = 0.5, pch = 1)
         panel.grid(h = -1, v = 2)
         panel.abline(v = 0, lty = 2)
         if (length(x) > 5) panel.loess(x, y, span = 0.85,
                                        col = 1, lwd = 2)
         })

# have a look at the consumption by treatment
xyplot(corrected_consumption ~ day | treatment, data = sucrose,
       xlab = "Day", ylab = "Sucrose Consumption",
       panel=function(x, y){
         panel.xyplot(x, y, col = 1, cex = 0.5, pch = 1)
         panel.grid(h = -1, v = 2)
         panel.abline(v = 0, lty = 2)
         if (length(x) > 5) panel.loess(x, y, span = 0.6,
                                        col = 1, lwd = 2)
         })

# have a look at the consumption by batch
xyplot(corrected_consumption ~ day | batch, data = sucrose,
       xlab = "Day", ylab = "Sucrose Consumption",
       panel=function(x, y){
         panel.xyplot(x, y, col = 1, cex = 0.5, pch = 1)
         panel.grid(h = -1, v = 2)
         panel.abline(v = 0, lty = 2)
         if (length(x) > 5) panel.loess(x, y, span = 0.6,
                                        col = 1, lwd = 2)
         })

# have a look at the consumption by treatment*batch
xyplot(corrected_consumption ~ day | batch*treatment, data = sucrose,
       xlab = "Day", ylab = "Sucrose Consumption",
       panel=function(x, y){
         panel.xyplot(x, y, col = 1, cex = 0.5, pch = 1)
         panel.grid(h = -1, v = 2)
         panel.abline(v = 0, lty = 2)
         if (length(x) > 5) panel.loess(x, y, span = 0.6,
                                        col = 1, lwd = 2)
         })

```

First determine the random effects component. Use beyond optimal model:

Doesn't look like variance increases with day but overall variance may be slightly different between treatment*batch combinations

Not interested in the effect of microcolony so should include this as a random effect. Intercept and slope of both seem to change between microcolonies.

```{r}
# beyond optimal model.

# biologically this is my beyond optimal model. I can imagine both treatment and batch potentially affecting the 
# starting (intercept) consumption and the relationship between consumption and day. 

M.lm <- lm(corrected_consumption ~ day + treatment + batch + treatment:batch + treatment:day + batch:day, data = sucrose)

summary(M.lm)

# use the residuals from this beyond optimal model
plot(M.lm, which = c (1))
```

Violation of heterogeneity. Let's start by determining our optimal random effects structure. 

```{r}
library(nlme)

M1.gls <- gls(corrected_consumption ~ day + treatment + batch + treatment:batch + treatment:day + batch:day, data = sucrose, method = "REML")

E <- resid(M1.gls)

op <- par(mfrow = c(2, 2))

boxplot(E ~ sucrose$treatment, main = "Treatment")
abline(0, 0)

boxplot(E ~ sucrose$batch, main = "Batch")
abline(0, 0)

boxplot(E ~ sucrose$treatment * sucrose$batch, main = "Treatment & Batch")
abline(0, 0)

boxplot(E ~ sucrose$microcolony_number, main = "Microcolony")
abline(0, 0)

par(op)

plot(E ~ sucrose$day, main = "Day")
abline(0, 0)

plot(M1.gls, which = c (1))
```

Day 2 batch3:control there are 4 points that are not in keeping with the rest of the data. Looks like I missed some slow leak spills. Without them my heterogeneity effectively disappears. 

Perform the analysis with and without them to show their effect on the process. 

Without them first. 

```{r}
sucrose <- sucrose [-c(105, 101, 107, 98),]

M1.gls <- gls(corrected_consumption ~ day + treatment + batch + treatment:batch + treatment:day + batch:day, data = sucrose, method = "REML")

E <- resid(M1.gls)

op <- par(mfrow = c(2, 2))

boxplot(E ~ sucrose$treatment, main = "Treatment")
abline(0, 0)

boxplot(E ~ sucrose$batch, main = "Batch")
abline(0, 0)

boxplot(E ~ sucrose$treatment * sucrose$batch, main = "Treatment & Batch")
abline(0, 0)

boxplot(E ~ sucrose$microcolony_number, main = "Microcolony")
abline(0, 0)

par(op)

plot(E ~ sucrose$day, main = "Day")
abline(0, 0)

plot(M1.gls, which = c (1))
```

Select random effect components. I think varident with treatment and batch may be a good idea and microcolony with a random intercept and slope. 

```{r}
# random intercept?
M1.lme <- lme(corrected_consumption ~ day + treatment + batch + 
                treatment:batch + treatment:day + 
                batch:day, data = sucrose, 
              random =~1 | microcolony_number, method = "REML")

AIC(M1.gls, M1.lme)

anova(M1.gls, M1.lme)

library(lmtest)
lrtest(M1.gls, M1.lme)

# random intercept is great improvement. Random intercept and slope?
M2.lme <-  lme(corrected_consumption ~ day + treatment + batch + treatment:batch + treatment:day + batch:day, 
               data = sucrose,
               random =~1 + day | microcolony_number, method = "REML")

AIC(M1.gls, M1.lme, M2.lme)

# lower aic for random intercept and slope model. Not a large improvement but still an improvement

# What about the variance structures?
M2a.lme <- lme(corrected_consumption ~ day + treatment + batch + treatment:batch + treatment:day + batch:day,
               data = sucrose,
               random =~1 + day | microcolony_number, method = "REML",
               weights = varIdent(form =~ 1 | treatment*batch))

M2b.lme <- lme(corrected_consumption ~ day + treatment + batch + treatment:batch + treatment:day + batch:day,
               data = sucrose,
               random =~1 + day | microcolony_number, method = "REML",
               weights = varIdent(form =~ 1 | batch))

M2c.lme <- lme(corrected_consumption ~ day + treatment + batch + treatment:batch + treatment:day + batch:day,
               data = sucrose,
               random =~1 + day | microcolony_number, method = "REML",
               weights = varIdent(form =~ 1 | treatment))

AIC(M2.lme, M2a.lme, M2b.lme, M2c.lme)

# batch varident is a significant improvement again. 
```

The optimal random effects structure is a random intercept and slope (1 + day | microcolony) with a variance structure that allows for a difference in variance between batch level.

Have a look at the residual plots of model with selected random effects structure. 

```{r}
M2b.lme <- lme(corrected_consumption ~ day + treatment + batch + treatment:batch + treatment:day + batch:day,
               data = sucrose,
               random =~1 + day | microcolony_number, method = "REML",
               weights = varIdent(form =~ 1 | batch))

# extract residuals
E.M2b.lme <- resid(M2b.lme, type = "normalized")

# residuals vs fitted
plot(M2b.lme, which = c (1))

# residuals vs day
plot(E.M2b.lme ~ sucrose$day, main = "Microcolony")
abline(0, 0)

op <- par(mfrow = c(2, 2))

boxplot(E.M2b.lme ~ sucrose$treatment, main = "Treatment")
abline(0, 0)

boxplot(E.M2b.lme ~ sucrose$batch, main = "Batch")
abline(0, 0)

boxplot(E.M2b.lme ~ sucrose$treatment * sucrose$batch, main = "Treatment & Batch")
abline(0, 0)

boxplot(E.M2b.lme ~ sucrose$microcolony_number, main = "Microcolony")
abline(0, 0)

par(op)
```

```{r}
# quick summary of model with optimal random effects structure
summary(M2b.lme)
```

Switch back to ML for fixed effects and start model selection 

```{r}
# set method to ML
M2b.lme <- lme(corrected_consumption ~ day + treatment + batch + treatment:batch + treatment:day + batch:day,
               data = sucrose,
               random =~1 + day | microcolony_number, method = "ML",
               weights = varIdent(form =~ 1 | batch))

# what does AIC say?
library(MASS)
stepAIC(M2b.lme)
```

Now manually

```{r}
# round 1
M2b.lme.a <- update(M2b.lme, .~. - treatment:batch)
M2b.lme.b <- update(M2b.lme, .~. - treatment:day)
M2b.lme.c <- update(M2b.lme, .~. - batch:day)

anova(M2b.lme, M2b.lme.a)
anova(M2b.lme, M2b.lme.b)
anova(M2b.lme, M2b.lme.c)

# remove treatment:batch

# round 2
M2b.lme.d <- update(M2b.lme.a, .~. - treatment:day)
M2b.lme.e <- update(M2b.lme.a, .~. - batch:day)

anova(M2b.lme.a, M2b.lme.d)
anova(M2b.lme.a, M2b.lme.e)

# remove treatment:day

# round 3
M2b.lme.f <- update(M2b.lme.d, .~. - treatment)
M2b.lme.g <- update(M2b.lme.d, .~. - batch:day)

anova(M2b.lme.d, M2b.lme.f)
anova(M2b.lme.d, M2b.lme.g)

# final model M2b.lme.d
```

```{r}
# refit with REML
M2b.lme.d <- update(M2b.lme.d, method = "REML")

summary(M2b.lme.d)
```

Model Validation 

```{r}
# extract residuals
E.M2b.lme.d <- resid(M2b.lme.d, type = "pearson")

# residuals vs fitted
plot(M2b.lme.d, which = c (1))

# residuals vs day
xyplot(E.M2b.lme.d ~ day, data = sucrose,
       xlab = "Day", ylab = "Pearson Residuals",
       panel=function(x, y){
         panel.xyplot(x, y, col = 1, cex = 0.5, pch = 1)
         panel.grid(h = -1, v = 2)
         panel.abline(v = 0, lty = 2)
         if (length(x) > 5) panel.loess(x, y, span = 0.6,
                                        col = 1, lwd = 2)
         })

# batches are different
xyplot(E.M2b.lme.d ~ day | batch, data = sucrose,
       xlab = "Day", ylab = "Pearson Residuals",
       panel=function(x, y){
         panel.xyplot(x, y, col = 1, cex = 0.5, pch = 1)
         panel.grid(h = -1, v = 2)
         panel.abline(v = 0, lty = 2)
         if (length(x) > 5) panel.loess(x, y, span = 0.6,
                                        col = 1, lwd = 2)
         })

xyplot(E.M2b.lme.d ~ day | batch*treatment, data = sucrose,
       xlab = "Day", ylab = "Pearson Residuals",
       panel=function(x, y){
         panel.xyplot(x, y, col = 1, cex = 0.5, pch = 1)
         panel.grid(h = -1, v = 2)
         panel.abline(v = 0, lty = 2)
         if (length(x) > 5) panel.loess(x, y, span = 0.6,
                                        col = 1, lwd = 2)
         })
```

Look at the visualisations

```{r}
# plot
# predict values
consumption <- predict(M2b.lme.d, value = "response")

# generate line components
dayvalues <- seq(1, 10, 0.1)

data_list <- list()

for (i in 1:length(unique(sucrose$batch))) {
  mydata <- data.frame(
    day = dayvalues,
    batch = unique(sucrose$batch) [i],
    treatment = unique(sucrose$treatment) [1])
  
  data_list [[i]] <- mydata
}

for (i in 1:length(unique(sucrose$batch))) {
  mydata <- data.frame(
    day = dayvalues,
    batch = unique(sucrose$batch) [i],
    treatment = unique(sucrose$treatment) [2])
  
  data_list [[i+3]] <- mydata
}

# re.form = NA, if NA or ~0, include no random effects.

line_list <- list()

for (i in 1:length(data_list)) {
  
  line <- predict(M2b.lme.d, data_list [[i]], type = "response", level = 0)
  
  line_list [[i]] <- line
  
}
```

[using predict in ggplot](https://aosmith.rbind.io/2018/11/16/plot-fitted-lines/#extracting-predicted-values-with-predict)

[limitations of confidence intervals with lme](https://bbolker.github.io/mixedmodels-misc/glmmFAQ.html#is-the-likelihood-ratio-test-reliable-for-mixed-models)

```{r}
library(tidyverse)
# confidence intervals. This method has its drawbacks though.
# Note that none of the following approaches takes the uncertainty of the random effects parameters into account
# If you want to take RE parameter uncertainty into account, a Bayesian approach is probably the easiest way to do it.
library(bernr)

# b1con
b1_con_ci_l <- bolker_ci(M2b.lme.d, data_list [[1]], pred_int = FALSE, conf_level = 0.95) [6]
b1_con_ci_h <- bolker_ci(M2b.lme.d, data_list [[1]], pred_int = FALSE, conf_level = 0.95) [7]

# b2con
b2_con_ci_l <- bolker_ci(M2b.lme.d, data_list [[2]], pred_int = FALSE, conf_level = 0.95) [6]
b2_con_ci_h <- bolker_ci(M2b.lme.d, data_list [[2]], pred_int = FALSE, conf_level = 0.95) [7]

# b3con
b3_con_ci_l <- bolker_ci(M2b.lme.d, data_list [[3]], pred_int = FALSE, conf_level = 0.95) [6]
b3_con_ci_h <- bolker_ci(M2b.lme.d, data_list [[3]], pred_int = FALSE, conf_level = 0.95) [7]

# b1fpf
b1_fpf_ci_l <- bolker_ci(M2b.lme.d, data_list [[4]], pred_int = FALSE, conf_level = 0.95) [6]
b1_fpf_ci_h <- bolker_ci(M2b.lme.d, data_list [[4]], pred_int = FALSE, conf_level = 0.95) [7]

# b2fpf
b2_fpf_ci_l <- bolker_ci(M2b.lme.d, data_list [[5]], pred_int = FALSE, conf_level = 0.95) [6]
b2_fpf_ci_h <- bolker_ci(M2b.lme.d, data_list [[5]], pred_int = FALSE, conf_level = 0.95) [7]

# b3fpf
b3_fpf_ci_l <- bolker_ci(M2b.lme.d, data_list [[6]], pred_int = FALSE, conf_level = 0.95) [6]
b3_fpf_ci_h <- bolker_ci(M2b.lme.d, data_list [[6]], pred_int = FALSE, conf_level = 0.95) [7]

# collect all required input into dataframes
b1_con_df <- cbind(data_list [[1]], line_list [[1]], b1_con_ci_l, b1_con_ci_h )
colnames(b1_con_df) [4] <- "corrected_consumption"
b1_con_df$batch_treatment <- as.factor(paste0(b1_con_df$batch, "-", b1_con_df$treatment, sep = ""))
  
b2_con_df <- cbind(data_list [[2]], line_list [[2]], b2_con_ci_l, b2_con_ci_h )
colnames(b2_con_df) [4] <- "corrected_consumption"
b2_con_df$batch_treatment <- as.factor(paste0(b2_con_df$batch, "-", b2_con_df$treatment, sep = ""))

b3_con_df <- cbind(data_list [[3]], line_list [[3]], b3_con_ci_l, b3_con_ci_h )
colnames(b3_con_df) [4] <- "corrected_consumption"
b3_con_df$batch_treatment <- as.factor(paste0(b3_con_df$batch, "-", b3_con_df$treatment, sep = ""))

# collect all required input into dataframes
b1_fpf_df <- cbind(data_list [[4]], line_list [[4]], b1_fpf_ci_l, b1_fpf_ci_h )
colnames(b1_fpf_df) [4] <- "corrected_consumption"
b1_fpf_df$batch_treatment <- as.factor(paste0(b1_fpf_df$batch, "-", b1_fpf_df$treatment, sep = ""))

b2_fpf_df <- cbind(data_list [[5]], line_list [[5]], b2_fpf_ci_l, b2_fpf_ci_h )
colnames(b2_fpf_df) [4] <- "corrected_consumption"
b2_fpf_df$batch_treatment <- as.factor(paste0(b2_fpf_df$batch, "-", b2_fpf_df$treatment, sep = ""))

b3_fpf_df <- cbind(data_list [[6]], line_list [[6]], b3_fpf_ci_l, b3_fpf_ci_h )
colnames(b3_fpf_df) [4] <- "corrected_consumption"
b3_fpf_df$batch_treatment <- as.factor(paste0(b3_fpf_df$batch, "-", b3_fpf_df$treatment, sep = ""))
```

```{r}
sucrose$batch_treatment <- as.factor(paste0(sucrose$batch, "-", sucrose$treatment, sep = ""))

p1 <- ggplot(sucrose, aes(x = day, y = corrected_consumption, colour = as.factor(batch_treatment))) + 
  
  geom_point(data = sucrose,
             aes(x = day,
                 y = corrected_consumption,
                 colour = as.factor(batch_treatment)),
             alpha = 0.25,
             position  = position_dodge(width = 0.75),
             show.legend = FALSE) +
  
  theme_bw() +
  
  geom_line(data = b1_con_df, aes(y = b1_con_df$corrected_consumption), size = 1) +
  
  geom_ribbon(data = b1_con_df, aes(ymin=b1_con_df$ci_l, ymax=b1_con_df$ci_h, fill = batch_treatment, color = NULL), alpha=0.2, show.legend = FALSE) +
  
  
  geom_line(data = b2_con_df, aes(y = b2_con_df$corrected_consumption), size = 1) +
  
  geom_ribbon(data = b2_con_df, aes(ymin=b2_con_df$ci_l, ymax=b2_con_df$ci_h, fill = batch_treatment, color = NULL), alpha=0.2, show.legend = FALSE) +
  

  geom_line(data = b3_con_df, aes(y = b3_con_df$corrected_consumption), size = 1) +
  
  geom_ribbon(data = b3_con_df, aes(ymin=b3_con_df$ci_l, ymax=b3_con_df$ci_h, fill = batch_treatment, color = NULL), alpha=0.2, show.legend = FALSE) +
  
  
  geom_line(data = b1_fpf_df, aes(y = b1_fpf_df$corrected_consumption), size = 1) +
  
  geom_ribbon(data = b1_fpf_df, aes(ymin=b1_fpf_df$ci_l, ymax=b1_fpf_df$ci_h, fill = batch_treatment, color = NULL), alpha=0.2, show.legend = FALSE) +
  
  
  geom_line(data = b2_fpf_df, aes(y = b2_fpf_df$corrected_consumption), size = 1) +
  
  geom_ribbon(data = b2_fpf_df, aes(ymin=b2_fpf_df$ci_l, ymax=b2_fpf_df$ci_h, fill = batch_treatment, color = NULL), alpha=0.2, show.legend = FALSE) +
  
  
  geom_line(data = b3_fpf_df, aes(y = b3_fpf_df$corrected_consumption), size = 1) +
  
  geom_ribbon(data = b3_fpf_df, aes(ymin=b3_fpf_df$ci_l, ymax=b3_fpf_df$ci_h, fill = batch_treatment, color = NULL), alpha=0.2, show.legend = FALSE) +
  
  

  labs(y = "Sucrose Consumption", colour = "Batch") +
  
  ggtitle("Sucrose Consumption During Exposure Window") +
    
  theme(plot.title = element_text(hjust = 0.5)) +
  
  scale_color_manual(values = c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2")) +
  
  scale_fill_manual(values = c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2")) +

  scale_x_continuous(breaks=seq(0,10,1)) +
  
  theme(legend.key = element_blank())

p1
```

Without ribbons

```{r}
p2 <- ggplot(sucrose, aes(x = day, y = corrected_consumption, colour = as.factor(batch_treatment))) + 
  
  geom_point(data = sucrose,
             aes(x = day,
                 y = corrected_consumption,
                 colour = as.factor(batch_treatment)),
             alpha = 0.25,
             position  = position_dodge(width = 0.75),
             show.legend = FALSE) +
  
  theme_bw() +
  
  geom_line(data = b1_con_df, aes(y = b1_con_df$corrected_consumption), size = 1) +
  
  geom_line(data = b2_con_df, aes(y = b2_con_df$corrected_consumption), size = 1) +
  
  geom_line(data = b3_con_df, aes(y = b3_con_df$corrected_consumption), size = 1) +
  
  geom_line(data = b1_fpf_df, aes(y = b1_fpf_df$corrected_consumption), size = 1) +
  
  geom_line(data = b2_fpf_df, aes(y = b2_fpf_df$corrected_consumption), size = 1) +
  
  geom_line(data = b3_fpf_df, aes(y = b3_fpf_df$corrected_consumption), size = 1) +
  
  labs(y = "Sucrose Consumption", colour = "Batch") +
  
  ggtitle("Sucrose Consumption During Exposure Window") +
    
  theme(plot.title = element_text(hjust = 0.5)) +
  
  scale_color_manual(values = c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2")) +
  
  scale_fill_manual(values = c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2")) +

  scale_x_continuous(breaks=seq(0,10,1)) +
  
  theme(legend.key = element_blank())

p2
```

The form of analysis I can perform is severely limited by the number of unique day points I have (6). This model says fpf group sucrose consumption at the beginning of the experiment was significantly higher. Then the pattern of sucrose consumption over time depended on batch. Batch 2 had decreasing sucrose consumption, which is in keeping with batch 2 issues during transportation. 

Of course these findings are undermined by patterns in the residuals. As there are only 6 timepoints, however, linear regression is really my only option. Or a really high span on a lowess. Then with lowess there is no way to perform mixed modelling. Also, with lowess the issue of having 6 time points is still present. 

Looking at residuals vs fitted for batch*treatment day panels. Seems to be a tendency for the fpf intercept to be too high (more negative residuals than positive, therefore predicted larger than actual). 

Due to the limits of this dataset I think it would be fair to reanalyse the dataset removing day and just looking at total sucrose consumption from day 1-10 for each treatment and batch. 

