---
title: "sucrose_consumption_aggregated"
author: "Guy Mercer"
date: "24/02/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r}
sucrose <- read.csv("output/sucrose_consumption_corrected_long.csv")
pollen <- read.csv("output/pollen_consumption_corrected_long.csv")
```

```{r}
library(tidyverse)

# x1000 for mg
sucrose$corrected_consumption <- sucrose$corrected_consumption * 1000
```

First correct day consumption values for worker number. 

```{r}
worker_number <- read.csv("input/number_of_workers_each_day.csv")

# convert into long format
day <- c("1", "2", "4", "6", "8", "10")

worker_long <- tibble()

for (i in 1:length(day)) {
  
  worker_temp <- worker_number [, c(1:5, 5+i)]
  
  worker_temp$day <- day [i]
  
  colnames(worker_temp) [6] <- "worker_number"
  
  worker_long <- rbind(worker_long, worker_temp)
  
}

# create microcolony_day variable for worker_long and sucrose
worker_long$mc_day <- paste0(worker_long$microcolony.number, "_", worker_long$day)

sucrose$mc_day <- paste0(sucrose$microcolony_number, "_", sucrose$day)

# associate the worker number with the sucrose dataframe. below says for each row of sucrose df
# scan each row of the worker_long df. if there is a match between the mc_day unique identifier
# then put the worker number value for that row of worker_long into the matched row of sucrose df
sucrose$worker_number <- 0

for (i in 1:nrow(sucrose)) {
  
  for (k in 1:nrow(worker_long)) {
    
      if (sucrose$mc_day [i] == worker_long$mc_day [k]) {sucrose$worker_number [i] <- worker_long$worker_number [k]}
    
  }
}

# remove mc_day from sucrose
sucrose <- sucrose [,-7]

# divide corrected consumption values by number of bees present
sucrose$corrected_consumption <- sucrose$corrected_consumption / sucrose$worker_number
```

Calculate average total consumption per bee per microcolony. May be more accurate to describe this as total sucrose processed per bee per microcolony.

```{r}
# sum consumption by MC
tot_con <- tibble()

microcolony_numbers <- unique(sucrose$microcolony_number)

for (i in 1:length(microcolony_numbers)) {
  
  if (length(sucrose$corrected_consumption [sucrose$microcolony_number == microcolony_numbers [i]]) == 6) {

  filtered_con <- cbind(sum(sucrose$corrected_consumption [sucrose$microcolony_number == microcolony_numbers [i]]), microcolony_numbers [i])
  
  tot_con <- rbind(tot_con, filtered_con)
  
  }

}

# re-associate these MC values with colony, batch etc information
batch_col_info <- unique(sucrose [, c(1,2,3,4)])

batch_col_info$corrected_consumption <- 0 

for (i in 1:nrow(tot_con)) {
  
  for (k in 1:nrow(batch_col_info)) {
    
    if (tot_con [i, 2] == batch_col_info [k, 1]) {
      
      batch_col_info$corrected_consumption [k] <- tot_con [i, 1] 
      
    }
    
  }
  
}

# remove 0s (these had spills) and rename data
sucrose <- batch_col_info

sucrose <- sucrose [sucrose$corrected_consumption > 0, ]

```

```{r}
# sort out the variable classes
sucrose$microcolony_number <- as.factor(sucrose$microcolony_number)
sucrose$batch <- as.factor(sucrose$batch)
sucrose$colony_derived_from <- as.factor(sucrose$colony_derived_from)
sucrose$treatment <- as.factor(sucrose$treatment)
```

Now begin analysis. At this point I have corrected the consumption values to account for number of bees present on particular day, summed those values and removed any MCs that didn't have full consumption dataset. 

Bit of exploratory analysis first

```{r}
# Cleveland Dot Plot. First make column with 1 and 2 that correspond to Treatment levels

sucrose$clevelandcode <- factor(
ifelse(sucrose$treatment == "control",
1,
2))

sucrose$clevelandcode <- as.numeric(sucrose$clevelandcode)

dotchart(sucrose$corrected_consumption,
         groups = factor(sucrose$clevelandcode),
         ylab = "Order of observations",
         xlab = "Consumption (mg)", main = "Cleveland dotplot", pch = sucrose$clevelandcode)

boxplot(corrected_consumption ~ treatment * batch,
        varwidth = TRUE, xlab = "TreatmentxBatch",
        main = "Boxplot of Consumption Conditional on Treatment and Batch", 
        ylab = "Consumption", data = sucrose)

boxplot(corrected_consumption ~ batch,
        varwidth = TRUE, xlab = "Batch",
        main = "Boxplot of Consumption Conditional on Batch", 
        ylab = "Consumption", data = sucrose)

boxplot(corrected_consumption ~ treatment,
        varwidth = TRUE, xlab = "Treatment",
        main = "Boxplot of Consumption Conditional on Treatment", 
        ylab = "Consumption", data = sucrose)
```

Looks like variance varies by treatment and colony combination. So include this as varIdent term in random structure. Colony is a potential random effect. Let's fit beyond optimal model. 

```{r}
# beyond optimal model.

# biologically this is my beyond optimal model.

M.lm <- lm(corrected_consumption ~ treatment + batch + treatment:batch, data = sucrose)

summary(M.lm)

# use the residuals from this beyond optimal model
plot(M.lm, which = c (1))
```

```{r}
library(nlme)

M1.gls <- gls(corrected_consumption ~ treatment + batch + treatment:batch, data = sucrose, method = "REML")

E <- resid(M1.gls)

op <- par(mfrow = c(2, 2))

boxplot(E ~ sucrose$treatment, main = "Treatment")
abline(0, 0)

boxplot(E ~ sucrose$batch, main = "Batch")
abline(0, 0)

boxplot(E ~ sucrose$treatment * sucrose$batch, main = "Treatment & Batch")
abline(0, 0)

boxplot(E ~ sucrose$colony_derived_from, main = "Colony")
abline(0, 0)

par(op)

plot(M1.gls, which = c (1))
```

Indicates colony should be included as a random effect. Batch seems to be driver of heterogeneity.

```{r}
# random intercept?
M1.lme <- lme(corrected_consumption ~ treatment + batch + treatment:batch, data = sucrose, 
              random = ~1 | colony_derived_from, method = "REML")

AIC(M1.gls, M1.lme)

anova(M1.gls, M1.lme)

library(lmtest)
lrtest(M1.gls, M1.lme)

# What about the variance structure?
M1a.gls <- gls(corrected_consumption ~ treatment + batch + treatment:batch,
               data = sucrose,
               method = "REML",
               weights = varIdent(form =~ 1 | treatment*batch))

M1b.gls <- gls(corrected_consumption ~ treatment + batch + treatment:batch,
               data = sucrose,
               method = "REML",
               weights = varIdent(form =~ 1 | batch))

M1c.gls <- gls(corrected_consumption ~ treatment + batch + treatment:batch,
               data = sucrose,
               method = "REML",
               weights = varIdent(form =~ 1 | treatment))

AIC(M1.gls, M1a.gls, M1b.gls, M1c.gls)

anova(M1.gls, M1b.gls)

# no varIdent structure is a significant improvement
```

Neither colony as a random effect or various variance structures are significant. So my starting model for model selection is M1.gls

```{r}
M1.gls <- gls(corrected_consumption ~ treatment + batch + treatment:batch, data = sucrose, method = "ML")

# what does AIC say?
library(MASS)
stepAIC(M1.gls)
```

Model: corrected_consumption ~ treatment + batch 

```{r}
# manual 

# round 1
M1.gls.a <- update(M1.gls, .~. - treatment:batch)

anova(M1.gls, M1.gls.a)

# round 2 
M1.gls.b <- update(M1.gls.a, .~. - treatment)
M1.gls.c <- update(M1.gls.a, .~. - batch)

anova(M1.gls.a, M1.gls.b)
anova(M1.gls.a, M1.gls.c)

```

Manually only batch has an effect. Refit with REML and look at validation plots. Zuur has a value of 0.04 in one of his examples and states in a series of hypothesis tests it is hardly convincing.  Page 113 of pdf document and page 297 with example of 0.026. Are the p values from the above approximate? Considering multiple hypothesis testing batch is questionable. Retain for now.

**All this is immaterial as treatment and batch are both part of the experimental design.**

```{r}
# refit with REML
M1.gls.a <- update(M1.gls.a, method = "REML")

summary(M1.gls.a)
```

```{r}
# extract residuals
E <- resid(M1.gls.a, type = "pearson")

# residuals vs fitted
plot(M1.gls.a, which = c (1))

op <- par(mfrow = c(2, 2))

boxplot(E ~ sucrose$treatment, main = "Treatment")
abline(0, 0)

boxplot(E ~ sucrose$batch, main = "Batch")
abline(0, 0)

boxplot(E ~ sucrose$treatment * sucrose$batch, main = "Treatment & Batch")
abline(0, 0)

boxplot(E ~ sucrose$colony_derived_from, main = "Colony")
abline(0, 0)

par(op)
```

As there are still patterns of heterogeneity for batch and colony. Add random and variance structures to the model and see if they improve the model validation plots. 

```{r}
# variance structures
M2a.lme <- lme(corrected_consumption ~ batch + treatment,
               data = sucrose,
               random =~1 | colony_derived_from, method = "REML",
               weights = varIdent(form =~ 1 | batch*treatment))

# extract residuals
E <- resid(M2a.lme, type = "normalized")

# residuals vs fitted
plot(M2a.lme, which = c (1))

op <- par(mfrow = c(2, 2))

boxplot(E ~ sucrose$treatment, main = "Treatment")
abline(0, 0)

boxplot(E ~ sucrose$batch, main = "Batch")
abline(0, 0)

boxplot(E ~ sucrose$treatment * sucrose$batch, main = "Treatment & Batch")
abline(0, 0)

boxplot(E ~ sucrose$colony_derived_from, main = "Colony")
abline(0, 0)

par(op)
```

Heterogeneity looks a lot better if colony is included as random intercept and batch*treatment as a variance structure. 

```{r}
# manual 
M3a.lme <- lme(corrected_consumption ~ batch + treatment + treatment:batch,
               data = sucrose,
               random =~1 | colony_derived_from, method = "ML",
               weights = varIdent(form =~ 1 | batch*treatment))

# round 1
M3a.lme.a <- update(M3a.lme, .~. - treatment:batch)

anova(M3a.lme, M3a.lme.a)

# round 2 
M3a.lme.b <- update(M3a.lme.a, .~. - treatment)
M3a.lme.c <- update(M3a.lme.a, .~. - batch)

anova(M3a.lme.a, M3a.lme.b)
anova(M3a.lme.a, M3a.lme.c)

```

Adding in the random and variance structures, which are both required for heterogeneity and due to experimental design, means nothing is significant.

M3a.lme.a final model. Interaction term was not significant.

Standard error extracted using these sources

[ref 1](https://stat.ethz.ch/pipermail/r-sig-mixed-models/2012q1/017530.html)
[ref 2](https://stackoverflow.com/questions/14358811/extract-prediction-band-from-lme-fit)

Only based on fixed component of the model.

```{r}
M3a.lme.a <- lme(corrected_consumption ~ batch + treatment,
               data = sucrose,
               random =~1 | colony_derived_from, method = "REML",
               weights = varIdent(form =~ 1 | batch*treatment))

# summary(M3a.lme.a)

newdat <- expand.grid(treatment=levels(sucrose$treatment), batch=levels(sucrose$batch))

newdat$pred <- predict(M3a.lme.a, newdat, level = 0)

Designmat <- model.matrix(eval(eval(M3a.lme.a$call$fixed)[-2]), newdat[-3])

predvar <- diag(Designmat %*% M3a.lme.a$varFix %*% t(Designmat))

newdat$SE <- sqrt(predvar)

newdat$ci_lower <- newdat$pred - (1.96 * newdat$SE) # 1.96 or qt(1 - (0.05 / 2), n - 1)?

newdat$ci_higher <- newdat$pred + (1.96 * newdat$SE) # 1.96 or qt(1 - (0.05 / 2), n - 1)?

batch_one_only <- newdat [newdat$batch == "1", ]

```

Treatment Plot

```{r}
# ggplot
p1 <- ggplot(data = batch_one_only, aes(x = treatment, y = pred, colour = as.factor(treatment))) + 
  
  geom_point(position = position_dodge(width = 0.75), size = 3) +
  
  geom_errorbar(aes(x = treatment,
                    ymin = ci_lower,
                    ymax = ci_higher,
                    colour = as.factor(treatment)),
                position = position_dodge(width = 0.75),
                width = 0.05, 
                size = 0.8) +
  
  geom_point(data = sucrose,
             aes(x = treatment,
                 y = corrected_consumption,
                 colour = as.factor(treatment)),
             alpha = 0.4,
             position  = position_jitterdodge(),
             show.legend = FALSE) +
  
  theme_bw() +

  labs(y = "Average Sucrose Consumption/Bee/Microcolony", colour = "Treatment") +
    
  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    
  ggtitle("Average Sucrose Consumption By Treatment") +
    
  theme(plot.title = element_text(hjust = 0.5), axis.title.x = element_blank()) +
  
  scale_color_manual(values = c("#117733", "#332288"), labels = c("Control", "Flupyradifurone")) +
  
  scale_x_discrete(labels=c("Control", "Flupyradifurone")) +
  
  annotate(geom = "text", x = 2.32, y = 4100, label = "bolditalic('p')~bold('= 0.248')", fontface = 2, size = 6, parse = TRUE)


p1
```

Now a batch plot for treatment=control

```{r}
control_only <- newdat [newdat$treatment == "control", ]

p <- ggplot(data = control_only, aes(x = batch, y = pred, colour = as.factor(batch))) + 
  
  geom_point(position = position_dodge(width = 0.75), size = 3) +
  
  geom_errorbar(aes(x = batch,
                    ymin = ci_lower,
                    ymax = ci_higher,
                    colour = as.factor(batch)),
                position = position_dodge(width = 0.75),
                width = 0.05, 
                size = 0.8) +
  
  geom_point(data = sucrose,
             aes(x = batch,
                 y = corrected_consumption,
                 colour = as.factor(batch)),
             alpha = 0.4,
             position  = position_jitterdodge(),
             show.legend = FALSE) +
  
  theme_bw() +

  labs(y = "Average Sucrose Consumption/Bee/Microcolony", colour = "Batch") +
    
  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    
  ggtitle("Average Sucrose Consumption By Batch") +
    
  theme(plot.title = element_text(hjust = 0.5), axis.title.x = element_blank()) +
  
  scale_color_manual(values = c("#88CCEE", "#CC6677", "#DDCC77")) +

  annotate(geom = "text", x = 3.25, y = 4100, label = "bolditalic('p')~bold('= 0.053')", fontface = 2, size = 6, parse = TRUE)

p
```

```{r}
pdf(file = "./output/sucrose_consumption_batch.pdf",   # The directory you want to save the file in
  width = 8, # The width of the plot in inches
  height = 6) # The height of the plot in inches

p

 dev.off()
 
pdf(file = "./output/sucrose_consumption_treatment.pdf",   # The directory you want to save the file in
  width = 8, # The width of the plot in inches
  height = 6) # The height of the plot in inches

p1

 dev.off()
```

