---
title: "male Number"
author: "Guy Mercer"
date: "21/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

Import data

```{r}
males <- read.csv("input/number_of_males.csv")

males$batch <- as.factor(males$batch)
males$colony <- as.factor(males$colony)
males$treatment <- as.factor(males$treatment)

males$number_of_males <- as.numeric(males$number_of_males)
```

Start with a frequency plot 

```{r}
plot(table(males$number_of_males))
```

Start with zero inflated poisson. Look at residuals vs fitted. Patterns? Compare to zero inflated NB with likelihood ratio test. Select appropriate. Patterns in residuals better? Plot against random effect variable. Move onto ZIGLMM using glmmTMB. Compare zero inflated model to zero inflated mixed model using a likelihood ratio test. Select appropriate. 

At this stage I will have appropriate distribution and random effect structure. Next, perform fixed effect model selection. Complete and check residuals vs fitted and residuals against remaining explanatory variables. All ok? 

```{r}
boxplot(number_of_males ~ treatment,
        varwidth = TRUE, xlab = "Treatment",
        main = "Boxplot of male Number Conditional on Treatment", 
        ylab = "Number of males", data = males)

males$clevelandcode <- factor(
ifelse(males$treatment == "control",
1,
2))

males$clevelandcode <- as.numeric(males$clevelandcode)


dotchart(males$number_of_males,
         groups = factor(males$clevelandcode),
         ylab = "Order of observations",
         xlab = "male Number", main = "Cleveland dotplot", pch = males$clevelandcode)
```

No outliers. Heterogeneity looks similar between treatment groups too. 

Fit ZIP. Only have a random intercept as none of my covariates will explain the probability of false zeros. False zeros were equally likely due to the experiment being cut short. Compare to ZINB.

```{r}
library(glmmTMB)

M1 <- glmmTMB(number_of_males ~ treatment * batch,
              data = males,
              ziformula=~1,
              family = poisson)

M2 <- glmmTMB(number_of_males ~ treatment * batch,
              data = males,
              ziformula=~1,
              family = "nbinom2",)

anova(M1,M2)
```

Negative binomial is still overdispersed deviance = 281.8 / df.residuals = 58 = 4.85. A quasi-negative binomial option does not exist. 

Likelihood ratio test shows ZINB is an improvement on the ZIP.

Take a look at the residuals. Refit with pscl as I can extract the residuals from this.

```{r}
library(pscl)

Zip1 <- zeroinfl(number_of_males ~ treatment * batch | 1, dist = "poisson",
                 link = "logit", data = males)

Zinb1 <- zeroinfl(number_of_males ~ treatment * batch | 1, dist = "negbin",
                 link = "logit", data = males)

library(lmtest)
lrtest(Zip1,Zinb1)

EP <- residuals(Zinb1, type = "pearson")

mu <- predict(Zinb1, type = "response")

plot(x = mu, y = EP, main = "Pearson Residuals")

boxplot(EP ~ males$colony, main = "Colony")
abline(0, 0)
```

Comparison of anova(M1,M2) and lrtest(Zip1,Zinb1) shows REML = FALSE for glmmTMB default. 

Residuals vs colony shows colony is important so fit as a random effect. 

```{r}
M3 <- glmmTMB(number_of_males ~ treatment * batch,
              data = males,
              ziformula=~1,
              family = "nbinom2",
              REML = TRUE)


M4 <- glmmTMB(number_of_males ~ treatment * batch + (1 | colony),
              data = males,
              ziformula=~1,
              family = "nbinom2",
              REML = TRUE)

# again I don't understand why lrtest isn't performing boundary correction. Divide by 2
# for random intercept vs no random intercept model. 
lrtest(M3, M4)
anova(M3, M4)
```

Here, if the testing the boundary correction is applicable it will take the p value under 0.05. Still approximate though and very close to threshold. Therefore, don't include colony as a random effect. 

Or perform analysis twice, with colony and without colony and see if there are any differences (doubt it)

```{r}
# with colony
M5 <- glmmTMB(number_of_males ~ treatment * batch + (1 | colony),
              data = males,
              ziformula=~1,
              family = "nbinom2")

M5a <- glmmTMB(number_of_males ~ treatment + batch + (1 | colony),
              data = males,
              ziformula=~1,
              family = "nbinom2")

anova(M5, M5a)

# removed interaction. M5a reference.

# remove batch
M5b <- glmmTMB(number_of_males ~ treatment + (1 | colony),
              data = males,
              ziformula=~1,
              family = "nbinom2")

anova(M5a, M5b)

# remove treatment
M5c <- glmmTMB(number_of_males ~ batch + (1 | colony),
              data = males,
              ziformula=~1,
              family = "nbinom2")

anova(M5a, M5c)

# treatment least significant so remove. M5c reference
M5d <- glmmTMB(number_of_males ~ 1 + (1 | colony),
              data = males,
              ziformula=~1,
              family = "nbinom2")

anova(M5c, M5d)

# batch not significant either. M5d final model. 
```

```{r}
summary(M5d)
```

Let's do it again without colony as a random intercept. 

```{r}
Zinb1a <- zeroinfl(number_of_males ~ treatment + batch | 1, dist = "negbin",
                 link = "logit", data = males)

lrtest(Zinb1, Zinb1a)

# Zinb1a reference
Zinb1b <- zeroinfl(number_of_males ~ treatment | 1, dist = "negbin",
                 link = "logit", data = males)

lrtest(Zinb1a, Zinb1b)

Zinb1c <- zeroinfl(number_of_males ~ batch | 1, dist = "negbin",
                 link = "logit", data = males)

lrtest(Zinb1a, Zinb1c)

# Zinb1c reference
Zinb1d <- zeroinfl(number_of_males ~ 1 | 1, dist = "negbin",
                 link = "logit", data = males)

lrtest(Zinb1c, Zinb1d)
```

```{r}
summary(Zinb1d)
```

Same conclusion that nothing is significant apart from the intercept. For simplicity remove colony as random effect. Model validation

```{r}
ED1 <- resid(Zinb1d, type = "pearson")

mu <- predict(Zinb1d, type = "response")

plot(x = mu, y = ED1, main = "Deviance residuals")

boxplot(ED1 ~ treatment,
        varwidth = TRUE, xlab = "Treatment",
        main = "Residuals vs Treatment", 
        ylab = "Residuals", data = males)

boxplot(ED1 ~ batch,
        varwidth = TRUE, xlab = "Batch",
        main = "Residuals vs Batch", 
        ylab = "Residuals", data = males)
```

For report, use these figures for batch, treatment and treatment:batch

```{r}
Zinb1a <- zeroinfl(number_of_males ~ treatment + batch | 1, dist = "negbin",
                 link = "logit", data = males)

lrtest(Zinb1, Zinb1a)

Zinb1b <- zeroinfl(number_of_males ~ treatment | 1, dist = "negbin",
                 link = "logit", data = males)

Zinb1c <- zeroinfl(number_of_males ~ batch | 1, dist = "negbin",
                 link = "logit", data = males)

Zinb1d <- zeroinfl(number_of_males ~ 1 | 1, dist = "negbin",
                 link = "logit", data = males)

lrtest(Zinb1c, Zinb1d)

lrtest(Zinb1b, Zinb1d)
```

Use this for colony

```{r}
M3 <- glmmTMB(number_of_males ~ treatment * batch,
              data = males,
              ziformula=~1,
              family = "nbinom2",
              REML = TRUE)


M4 <- glmmTMB(number_of_males ~ treatment * batch + (1 | colony),
              data = males,
              ziformula=~1,
              family = "nbinom2",
              REML = TRUE)

# again I don't understand why lrtest isn't performing boundary correction. Divide by 2
# for random intercept vs no random intercept model. 
lrtest(M3, M4)
anova(M3, M4)
```
 Refit with REML
 
```{r}
M3 <- glmmTMB(number_of_males ~ 1,
              data = males,
              ziformula=~1,
              family = "nbinom2",
              REML = TRUE)

summary(M3)
```
 
 Add treatment back in as that is what we are interested in. Want to show its lack of effect. 
 
```{r}
males$treatment <- relevel(males$treatment, "control")
M4 <- glmmTMB(number_of_males ~ 1 + treatment,
              data = males,
              ziformula=~1,
              family = "nbinom2",
              REML = TRUE)

summary(M4)

# control ci
con_ci <- confint(M4)

# fpf ci
males$treatment <- relevel(males$treatment, "fpf-pure")

M4 <- glmmTMB(number_of_males ~ 1 + treatment,
              data = males,
              ziformula=~1,
              family = "nbinom2",
              REML = TRUE)

fpf_ci <- confint(M4)

# assemble dataframe
row1 <- c(treatment="control", Prediction = exp(con_ci [1,3]), ci_l = exp(con_ci [1,1]), ci_h = exp(con_ci [1,2]))


row2 <- c(treatment="fpf-pure", Prediction = exp(fpf_ci [1,3]), ci_l = exp(fpf_ci [1,1]), ci_h = exp(fpf_ci [1,2]))

predictions <- as.data.frame(rbind(row1, row2))

predictions$Prediction <- as.numeric(as.character(predictions$Prediction))
predictions$ci_l <- as.numeric(as.character(predictions$ci_l))
predictions$ci_h <- as.numeric(as.character(predictions$ci_h))
```
 
```{r}
library(tidyverse)

# ggplot
p1 <- ggplot(data = predictions, aes(x = treatment, y = Prediction, colour = as.factor(treatment))) + 
  
  geom_point(position = position_dodge(width = 0.75), size = 3) +
  
  geom_errorbar(aes(x = treatment,
                    ymin = ci_l,
                    ymax = ci_h,
                    colour = as.factor(treatment)),
                position = position_dodge(width = 0.75),
                width = 0.05, 
                size = 0.8) +
  
  geom_point(data = males,
             aes(x = treatment,
                 y = number_of_males,
                 colour = as.factor(treatment)),
             alpha = 0.4,
             position  = position_jitterdodge(),
             show.legend = FALSE) +
  
  theme_bw() +

  labs(y = "Number of Males", colour = "Treatment") +
    
  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    
  ggtitle("Male Number") +
    
  theme(plot.title = element_text(hjust = 0.5), axis.title.x = element_blank()) +
  
  scale_color_manual(values = c("#117733", "#332288")) 

p1
```

```{r}
cairo_pdf(file = "./treatment_boxplot.pdf",   # The directory you want to save the file in
  width = 8, # The width of the plot in inches
  height = 6) # The height of the plot in inches

p1

 dev.off()
```


Report the probability of false zeros is exp(-0.699220735) / 1 + exp(-0.699220735). This is 0.33. 

Probabilities have to add to 1. The full range of outcomes is false zero, true zero and count. 

The probability that (yi = 0) = P(false zero) + (1-P(false zero) x P(true zero))

Let's unpack this. Probability of 0 is equal to the probability of a false zero plus the probability of an observation not being a false zero (everything else) multiplied by the probability of a true zero. 

Probability that (yi > 0) = 1-P(false zero) x P(true count)

so P(yi = 0) + P(yi > 0) = 1

So for my data above where the mean for both groups is approximately 6. The probability of a 0 from a poisson with a mean of 6 is equal to e^-6, which is 0.25%. Therefore, the probability of a zero = 0.33 + (1-0.33)x0.0025 = 0.3317. 

So if I calculate the probability of a false zero in terms of just the zeros  = 0.33/0.3317 = 0.995. 

So basically all my zeros are false zeros. 

The probability of false zeros therefore isn't showing the probability of a false zero in terms of just zeros, but in terms of of the whole data. 

Therefore, if the mean of a poisson is 6 and my P(false zero) = 0.33, P(true zero) = (1-0.33)xe^-6) = 0.0017, P(yi > 0) = (1-0.33)xPr(Count Process)

Pr(Count Process) = 1 - e^-6 = 0.9975 x (1-0.33) = 0.6683

Pr(total) = 0.33 + 0.0017 + 0.6683 = 1


Take home is the probability of false zeros is in terms of total probability, not only conditional on zero. 