---
title: "Male Mass Analysis"
author: "Guy Mercer"
date: "20/01/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

Load tidyverse

```{r}
library(tidyverse)
```

Import data and give all the variables the suitable class

```{r}
male_mass_df <- read.csv("input/male_mass.csv")

male_mass_df$male_number <- as.numeric(male_mass_df$male_number)

male_mass_df$microcolony_number <- as.factor(male_mass_df$microcolony_number)

male_mass_df$batch <- as.factor(male_mass_df$batch)

male_mass_df$colony_number <- as.factor(male_mass_df$colony_number)

male_mass_df$treatment <- as.factor(male_mass_df$treatment)

male_mass_df$dry_mass_g <- as.numeric(male_mass_df$dry_mass_g)
```

Express male mass in terms of mg not g and update column header

```{r}
male_mass_df$dry_mass_g <- male_mass_df$dry_mass_g * 1000

colnames(male_mass_df) [6] <- "male_mass_mg"
```

Data Exploration

```{r}
# Cleveland Dot Plot. First make column with 1 and 2 that correspond to Treatment levels

male_mass_df$clevelandcode <- factor(
ifelse(male_mass_df$treatment == "control",
1,
2))

male_mass_df$clevelandcode <- as.numeric(male_mass_df$clevelandcode)

dotchart(male_mass_df$male_mass_mg,
         groups = factor(male_mass_df$clevelandcode),
         ylab = "Order of observations",
         xlab = "male Mass (mg)", main = "Cleveland dotplot", pch = male_mass_df$clevelandcode)

boxplot(male_mass_mg ~ treatment * batch,
        varwidth = TRUE, xlab = "Treatment",
        main = "Boxplot of male Mass Conditional on Treatment and Batch", 
        ylab = "male Mass", data = male_mass_df)

boxplot(male_mass_mg ~ batch,
        varwidth = TRUE, xlab = "Treatment",
        main = "Boxplot of male Mass Conditional on Batch", 
        ylab = "male Mass", data = male_mass_df)

boxplot(male_mass_mg ~ treatment,
        varwidth = TRUE, xlab = "Treatment",
        main = "Boxplot of male Mass Conditional on Treatment", 
        ylab = "male Mass", data = male_mass_df)
```

Data Exploration wasn't difficult as I haven't got many explanatory variables to choose from. I think a suitable starting model is

male mass ~ treatment * batch + (1 | colony/microcolony) 

Let's begin with a linear model

```{r}
M.lm <- lm(male_mass_mg ~ treatment * batch,
           data = male_mass_df)

drop1(M.lm, test = "F")

# and the residuals vs fitted
plot(M.lm, which = c (1))
```

Violation of homogeneity. Refit with gls and look at the residuals vs treatment, batch and treatment*batch. Also look at residuals vs microcolony. 

```{r}
library(nlme)
M1.gls <- gls(male_mass_mg ~ treatment * batch,
           data = male_mass_df)

E <- resid(M1.gls)

op <- par(mfrow = c(2, 2))

boxplot(E ~ male_mass_df$treatment, main = "Treatment")
abline(0, 0)

boxplot(E ~ male_mass_df$batch, main = "Batch")
abline(0, 0)

boxplot(E ~ male_mass_df$treatment * male_mass_df$batch, main = "Treatment & Batch")
abline(0, 0)

boxplot(E ~ male_mass_df$microcolony_number, main = "Microcolony")
abline(0, 0)
```

The variance looks different for various treatment batch combinations. Also, for some microcolonies all the residuals are above or below 0, showing it should be included as an explanatory effect, in this case a random effect. 

Fit varIdent variance structure and microcolony/colony as a random effect. 

```{r}
# start with random intercept
M1.lme <- lme(male_mass_mg ~ treatment * batch,
              data = male_mass_df,
              random = ~ 1 | colony_number/microcolony_number, method = "REML")

AIC(M1.lme, M1.gls)
anova(M1.gls, M1.lme)

# large improvement. Add varIdent
M2.lme <- lme(male_mass_mg ~ treatment * batch,
              data = male_mass_df,
              weights = varIdent(form = ~ 1 | treatment * batch),
              random = ~ 1 | colony_number/microcolony_number, method = "REML")

anova(M1.lme, M2.lme)

# small improvement

# plot the standardised residuals versus fitted values
plot(M1.lme, col = 1) 
plot(M2.lme, col = 1) 

E1 <- resid(M1.lme, type = "normalized")
boxplot(E1 ~ male_mass_df$treatment * male_mass_df$batch, main = "Treatment & Batch")
abline(0, 0)

E2 <- resid(M2.lme, type = "normalized")
boxplot(E2 ~ male_mass_df$treatment * male_mass_df$batch, main = "Treatment & Batch")
abline(0, 0)
```

First point of slight confusion. Slightly different approaches are taken when finding the optimal random structure in the textbook. One example fits a random intercept and variance structure and then drops each of them from the model. Then compares anova(). If I do this both are significant. I drop the random intercept and p<0.0001. I drop the varIdent and p < 0.0306. What I have done above is analagous to this. 

If, however, I had started with the variance structure and compared to M1.gls, the basic model, none of the the varIdent options (treatment*batch, batch, treatment) were improvements. This is another common approach Zuur takes. 

I have looked at the residuals vs fitted for each model and residuals vs batchxtreatment and adding batchxtreatment variance structure does  seem to improve it. Therefore I am going to keep it.

Continue with M1.lme. Model selection of fixed terms.

```{r}
# redefine with ML
# I realise there is a naming issue here. M2.lme has become M1.lme. It has no effect on the analysis though. 
M1.lme <- lme(male_mass_mg ~ treatment * batch,
              data = male_mass_df,
              weights = varIdent(form = ~ 1 | treatment * batch),
              random = ~ 1 | colony_number/microcolony_number, method = "ML")

# drop interaction term
M1.lmeA <- update(M1.lme, .~. -batch:treatment)

anova(M1.lme, M1.lmeA)

# dropped interaction term

# drop batch
M1.lmeB <- update(M1.lmeA, .~. -batch)

anova(M1.lmeA, M1.lmeB)

# drop treatment 
M1.lmeC <- update(M1.lmeA, .~. -treatment)

anova(M1.lmeA, M1.lmeC)

# dropped treatment
M1.lmeD <- update(M1.lmeC, .~. -batch)

anova(M1.lmeC, M1.lmeD)

# nothing left bar an intercept

# add back in treatment as I am interested in its effect, or lack of. Model validation.
M1.lmeE <- lme(male_mass_mg ~ 1 + treatment,
               data = male_mass_df,
               random = ~ 1 | colony_number/microcolony_number, method = "ML")

plot(M1.lmeE, col = 1)

# refit model with REML
M1.lmeE <- update(M1.lmeE, method = "REML")

summary(M1.lmeE)
```

Visualisations

```{r}
boxplot(predict(M1.lmeE, level = 0, interval = confidence) ~ treatment, data = male_mass_df)
```

Refit with lmer to (more) easily get confidence intervals

```{r}
library(lme4)
# https://stackoverflow.com/questions/11072544/how-to-get-coefficients-and-their-confidence-intervals-in-mixed-effects-models

# bit of a nightmare returning the CIs. Have to specify model with lmer, run it, run confint, then relevel to get CIs for
# other treatment level. 
M1.lmer <- lmer(male_mass_mg ~ 1 + treatment + (1 | colony_number/microcolony_number), 
                data = male_mass_df, REML = TRUE)

M1.lmer.CI <- confint(M1.lmer)

M1.lmer.coeff <- summary(M1.lmer)$coefficients

control <- c(summary(M1.lmer)$coefficients [1], M1.lmer.CI [4, 1], M1.lmer.CI [4, 2])

male_mass_df$treatment <- relevel(male_mass_df$treatment, "fpf-pure")

M1.lmer <- lmer(male_mass_mg ~ 1 + treatment + (1 | colony_number/microcolony_number), 
                data = male_mass_df, REML = TRUE)

M1.lmer.CI <- confint(M1.lmer)

M1.lmer.coeff <- summary(M1.lmer)$coefficients

fpf_pure <- c(summary(M1.lmer)$coefficients [1], M1.lmer.CI [4, 1], M1.lmer.CI [4, 2])


# use this for the visualisation plus the raw data. 
M1.lmer.df <- data.frame(control, fpf_pure)
```

```{r}
# get df into right format
M1.lmer.df <- t(M1.lmer.df)
treatment <- c("control", "fpf-pure")
M1.lmer.df <- cbind(M1.lmer.df, treatment)
colnames(M1.lmer.df) <- c("prediction", "ci_l", "ci_h", "treatment")
rownames(M1.lmer.df) <- c()
M1.lmer.df <- as.data.frame(M1.lmer.df)
M1.lmer.df$treatment <- as.factor(M1.lmer.df$treatment)
M1.lmer.df$prediction <- as.numeric(as.character(M1.lmer.df$prediction))
M1.lmer.df$ci_l <- as.numeric(as.character(M1.lmer.df$ci_l))
M1.lmer.df$ci_h <- as.numeric(as.character(M1.lmer.df$ci_h))

# create df for adding p value to plot
group1 <- "control"
group2 <- "treatment"
# summary(M1.lmeE) [20]
p <- 6.792155e-01
p.signif <- "ns"
p_data <- as.data.frame(cbind(group1, group2, p , p.signif))
```

```{r}
# library(ggpubr)

# ggplot
p1 <- ggplot(data = M1.lmer.df, aes(x = treatment, y = prediction, colour = as.factor(treatment))) + 
  
  geom_point(position = position_dodge(width = 0.75), size = 3) +
  
  geom_errorbar(aes(x = treatment,
                    ymin = ci_l,
                    ymax = ci_h,
                    colour = as.factor(treatment)),
                position = position_dodge(width = 0.75),
                width = 0.05, 
                size = 0.8) +
  
  geom_point(data = male_mass_df,
             aes(x = treatment,
                 y = male_mass_mg,
                 colour = as.factor(treatment)),
             alpha = 0.25,
             position  = position_jitterdodge(),
             show.legend = FALSE) +
  
  theme_bw() +

  labs(y = "male Mass (mg)", colour = "Treatment") +
    
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
    
  ggtitle("male Mass") +
    
  theme(plot.title = element_text(hjust = 0.5), axis.title.x = element_blank()) +
  
  scale_color_manual(values = c("#117733", "#332288")) 

p1

# can't install ggpubr because quantreg package is playing up add this to get p values when I can 
# +
# stat_pvalue_manual(p_data, 
#                     y.position = 800, step.increase = 0.1,
#                     label = "p.adj.signif",
#                     size = 3,
#                     hide.ns = TRUE)

```


```{r}
cairo_pdf(file = "./treatment_boxplot.pdf",   # The directory you want to save the file in
  width = 8, # The width of the plot in inches
  height = 6) # The height of the plot in inches

p1

 dev.off()
```

